---
title: "Cal Poly Herpetology CURE - Baseline Analyses"
authors: "Savannah Weaver, Dylan Stephens, Jenna Heaphy"
date: "Spring Quarter 2021"
output: github_document
---

# Packages

```{r setup, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse")
if (!require("UsingR")) install.packages("UsingR")
library("UsingR")
```



# Data

### Morphometrics and Blood Data

variables:
- date
- collection/capture time for each lizard
- individual ID for each lizard
- sock ID used to capture each lizard
- SVL = snout-vent length
- mass in grams
- sex
- if female, whether or not gravid
- which eye the blood sample was taken from
- percent hematocrit = percent of blood that's red blood cells
- osmolality = a proxy of hydration, should be inversely related to water content of a lizard (this is the average of 1-3 replicates)
- cloacal temperature at the time of CEWL measurement
- processing time for each lizard, when all measurements were finished
- hemolyzed whether or not red blood cells burst and contaminated plasma

```{r morpho blood data}
morpho_blood_dat <- read.csv("./data/Herpetology_Data.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::mutate(date = as.Date(date, format = "%m/%d/%y"),
                collect_time = as.POSIXct(collect_time, format = "%H:%M"),
                processing_time = as.POSIXct(processing_time, format = "%H:%M"),
                sex_M_F = as.factor(sex_M_F),
                gravid_Y_N = as.factor(gravid_Y_N),
                blood_sample_eye = as.factor(blood_sample_eye),
                hemolyzed. = as.factor(hemolyzed.),
                individual_ID = as.character(individual_ID)
                ) %>%
  dplyr::filter(osmolality_mmol_kg <= 500) # remove one point I think inaccurate
summary(morpho_blood_dat)
```


### CEWL Data

```{r CEWL data}
# week 1
CEWL_April_05 <- read.csv("./data/4-5-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::mutate(week = "1") %>% # add weekend index for throughout season
  # I have to do this for each one 
  # so they all have the same number of columns for joining
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                CV = CV...., # rename
                SSWL_g_m2 = SSWL..g..m2.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH...., # rename
                week
                )

# week 2
CEWL_April_19 <- read.csv("./data/4-19-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::mutate(week = "2") %>% # add weekend index for throughout season
  # I have to do this for each one 
  # so they all have the same number of columns for joining
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                CV = CV...., # rename
                SSWL_g_m2 = SSWL..g..m2.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH...., # rename
                week
                )

# week 3
CEWL_April_26 <- read.csv("./data/4-26-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::mutate(week = "3") %>% # add weekend index for throughout season
  # I have to do this for each one 
  # so they all have the same number of columns for joining
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                CV = CV...., # rename
                SSWL_g_m2 = SSWL..g..m2.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH...., # rename
                week
                )

# merge all CEWL datafiles & reformat
CEWL <- CEWL_April_05 %>%
  # join
  rbind(., CEWL_April_19, CEWL_April_26) %>%
  # remove any unsuccessful measurements
  dplyr::filter(Status == "Normal") %>%
  # reformat data
  dplyr::mutate(date_time = as.POSIXct(paste(date, Time), format = "%m/%d/%y %I:%M:%S %p"),
                # date only
                date = as.Date(date, format = "%m/%d/%y"),
                # time with arbitrary date, need to figure out how to remove
                Time = as.POSIXct(Time, format = "%I:%M:%S %p"),
                # extract individual ID number
                individual_ID = as.numeric(substring(ID, 1, 2)),
                # extract body region for CEWL measurement
                region = as.factor(substring(ID, 4, 7)),
                week = as.factor(week),
                individual_ID = as.character(individual_ID)
                ) %>%
  dplyr::select(-ID, -Status
                ) %>%
  dplyr::filter(complete.cases(.))
summary(CEWL)
unique(CEWL$region)
```


### Split CEWL Data

```{r}
# select each CEWL region separately 
CEWL_dorsum <- CEWL %>%
  dplyr::filter(region == "dors") %>%
  dplyr::select(date, individual_ID, 
                dorsum_TEWL_g_m2h = TEWL_g_m2h)
CEWL_ventrum <- CEWL %>%
  dplyr::filter(region == "vent") %>%
  dplyr::select(date, individual_ID, 
                ventrum_TEWL_g_m2h = TEWL_g_m2h)
CEWL_dewlap <- CEWL %>%
  dplyr::filter(region == "dewl") %>%
  dplyr::select(date, individual_ID, 
                dewlap_TEWL_g_m2h = TEWL_g_m2h)
CEWL_head <- CEWL %>%
  dplyr::filter(region == "head") %>%
  dplyr::select(date, individual_ID, 
                head_TEWL_g_m2h = TEWL_g_m2h)
CEWL_mitepatch <- CEWL %>%
  dplyr::filter(region == "mite") %>%
  dplyr::select(date, individual_ID, 
                mitepatch_TEWL_g_m2h = TEWL_g_m2h)

# also get average across body regions
CEWL_avg <- CEWL %>%
  group_by(individual_ID) %>%
  summarise(avg_CEWL = mean(TEWL_g_m2h)) %>%
  dplyr::select(individual_ID, avg_CEWL)
```


### Rearrange and Join

```{r join data}
# join all CEWL regions to morpho and blood data
data_full <- morpho_blood_dat %>%
  left_join(CEWL_dorsum, by = c("date", "individual_ID")) %>%
  left_join(CEWL_ventrum, by = c("date", "individual_ID")) %>%
  left_join(CEWL_dewlap, by = c("date", "individual_ID")) %>%
  left_join(CEWL_head, by = c("date", "individual_ID")) %>%
  left_join(CEWL_mitepatch, by = c("date", "individual_ID")) %>%
  left_join(CEWL_avg, by = "individual_ID")
```


### Join Differently 

```{r}
CEWL_plus <- CEWL %>%
  left_join(morpho_blood_dat, by = c("date", "individual_ID")
            )
```




# Data Cleaning


## Histograms





# Basic Figures

## Hct ~ SVL

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = SVL_mm,
                 y = hematocrit_percent, 
                 #color = tmt
                 ), 
             size = 1, 
             alpha = 0.6) + 
  #scale_colour_manual(name = "Treatment", 
   #                   #labels = c("Burrow", "Open", "Shrub"), 
    #                  values = c("salmon1", "royalblue") ) +
  #scale_shape_discrete(name = "Microhabitat", 
   #                    labels = c("Burrow", "Open", "Shrub") ) + 
  stat_smooth(aes(x = SVL_mm, 
                  y = hematocrit_percent, 
                  #color = tmt
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("SVL") + 
  ylab("Hct") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
        #legend.position = c(0.9, 0.145)
)
```



## Osml ~ SVL

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = SVL_mm,
                 y = osmolality_mmol_kg, 
                 #color = tmt
                 ), 
             size = 1, 
             alpha = 0.6) + 
  #scale_colour_manual(name = "Treatment", 
   #                   #labels = c("Burrow", "Open", "Shrub"), 
    #                  values = c("salmon1", "royalblue") ) +
  #scale_shape_discrete(name = "Microhabitat", 
   #                    labels = c("Burrow", "Open", "Shrub") ) + 
  stat_smooth(aes(x = SVL_mm, 
                  y = osmolality_mmol_kg, 
                  #color = tmt
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("SVL") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
        #legend.position = c(0.9, 0.145)
)
```


## Hematocrit ~ Mass

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = mass_g,
                 y = hematocrit_percent, 
                 #color = tmt
                 ), 
             size = 1, 
             alpha = 0.6) + 
  #scale_colour_manual(name = "Treatment", 
   #                   #labels = c("Burrow", "Open", "Shrub"), 
    #                  values = c("salmon1", "royalblue") ) +
  #scale_shape_discrete(name = "Microhabitat", 
   #                    labels = c("Burrow", "Open", "Shrub") ) + 
  stat_smooth(aes(x = mass_g, 
                  y = hematocrit_percent, 
                  #color = tmt
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Mass") + 
  ylab("Hematocrit") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
        #legend.position = c(0.9, 0.145)
)
```

## Osml ~ Mass

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = mass_g,
                 y = osmolality_mmol_kg, 
                 #color = tmt
                 ), 
             size = 1, 
             alpha = 0.6) + 
  #scale_colour_manual(name = "Treatment", 
   #                   #labels = c("Burrow", "Open", "Shrub"), 
    #                  values = c("salmon1", "royalblue") ) +
  #scale_shape_discrete(name = "Microhabitat", 
   #                    labels = c("Burrow", "Open", "Shrub") ) + 
  stat_smooth(aes(x = mass_g, 
                  y = osmolality_mmol_kg), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "gray",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Mass") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
        #legend.position = c(0.9, 0.145)
)
```


## Hct ~ Sex

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = sex_M_F, 
                   y = hematocrit_percent, 
                   color = sex_M_F
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Sex", 
                      #labels = c("Shrubbed", "Shrubless"), 
                      values = c("green4", "salmon1") ) +
  theme_classic() + 
  xlab("Sex") + 
  ylab("Hematocrit") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)
```



## Osml ~ Sex

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = sex_M_F, 
                   y = osmolality_mmol_kg, 
                   color = sex_M_F
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Sex", 
                      #labels = c("Shrubbed", "Shrubless"), 
                      values = c("green4", "salmon1") ) +
  theme_classic() + 
  xlab("Sex") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)
```



# Important Figures


## CEWL ~ Body Region


```{r}
CEWL %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = region, 
                   y = TEWL_g_m2h, 
                   color = region
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_color_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  scale_x_discrete(labels = c("Dewlap", "Dorsum", "Head",
                              "Mite Patch", "Ventrum")) +
  theme_classic() + 
  xlab("Body Region") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)
```


## All CEWL ~ Osml

```{r}
CEWL_plus %>% 
  ggplot(data = .) + 
  geom_point(aes(x = osmolality_mmol_kg,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Osmolality") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
        #legend.position = c(0.9, 0.145)
)
```

## Dorsum CEWL ~ Osmolality 

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = osmolality_mmol_kg,
                 y = dorsum_TEWL_g_m2h, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = dorsum_TEWL_g_m2h, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1, 
              color = "lightblue4") + 
  theme_classic() + 
  xlab("Osmolality") + 
  ylab("Dorsum CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0)
```



## Ventrum CEWL ~ Osmolality 

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = osmolality_mmol_kg,
                 y = ventrum_TEWL_g_m2h, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = ventrum_TEWL_g_m2h, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1, 
              color = "seashell4") + 
  theme_classic() + 
  xlab("Osmolality") + 
  ylab("Ventrum CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0)
```



## Dewlap CEWL ~ Osmolality 

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = osmolality_mmol_kg,
                 y = dewlap_TEWL_g_m2h, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = dewlap_TEWL_g_m2h, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1, 
              color = "palegreen4") + 
  theme_classic() + 
  xlab("Osmolality") + 
  ylab("Dewlap CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0)
```



## Mite Patch CEWL ~ Osmolality 

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = osmolality_mmol_kg,
                 y = mitepatch_TEWL_g_m2h, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = mitepatch_TEWL_g_m2h, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1, 
              color = "lightpink3") + 
  theme_classic() + 
  xlab("Osmolality") + 
  ylab("Mite Patch CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0)
```


## Head CEWL ~ Osmolality 

```{r}
data_full %>% 
  ggplot(data = .) + 
  geom_point(aes(x = osmolality_mmol_kg,
                 y = head_TEWL_g_m2h, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = head_TEWL_g_m2h, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1, 
              color = "plum4") + 
  theme_classic() + 
  xlab("Osmolality") + 
  ylab("Head CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0)
```


## CEWL ~ Week

```{r}
CEWL %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = week, 
                   y = TEWL_g_m2h, 
                   color = week
                   ), 
               size = 1,
               alpha = 0.6) + 
  facet_wrap(~region) + # could not figure out how to change facet labels without changing underlying data
  scale_color_manual(values = c("royalblue1", "mediumorchid", "seagreen4")) +
  scale_x_discrete(breaks = c(1,2,3)) +
  theme_classic() + 
  xlab("Week") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)
```




# Weather Data






# Models

when2meet


   
# PCA













# Ideas






