---
title: "Cal Poly Herpetology CURE - Capture Data Analyses"
author: "Savannah Weaver"
date: "June 2021"
output: 
  pdf_document:
    toc: TRUE
---


# Packages

```{r setup, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse") # workflow and plots
if (!require("zoo")) install.packages("zoo")
library("zoo") # interpolation using na.approx
if (!require("weathermetrics")) install.packages("weathermetrics")
library("weathermetrics") # F to C conversion
if (!require("PerformanceAnalytics")) install.packages("PerformanceAnalytics") 
library("PerformanceAnalytics") # pretty multicollinearity plots
if (!require("UsingR")) install.packages("UsingR")
library("UsingR")
if (!require("MASS")) install.packages("MASS")
library("MASS") # model selection & box-cox
if (!require("lme4")) install.packages("lme4")
library("lme4") # for LMMs
if (!require("onewaytests")) install.packages("onewaytests")
library("onewaytests") # for Brown-Forsythe test
```


# Background and Goals

This data was collected April - May 2021 during a course-based undergraduate research experience (CURE) in Herpetology, taught by Emily Taylor at Cal Poly, San Luis Obispo. This part of the study was conducted to describe the variation in hydrophysiology in *Sceloporus occidentalis* and to investigate that drives that variation. Please refer to **doi:** for full details.

In this document, we investigate differences in cutaneous evaporative water loss (CEWL) across body regions and dependent on environment, body size, health, and hydration. 


# Data

### Morphometrics and Blood Data

This data was collected upon capture of each lizard.

Variables in this dataframe:
- date
- collection/capture time for each lizard
- individual ID for each lizard
- sock ID used to capture each lizard (removed, not relevant to analyses)
- SVL = snout-vent length
- mass in grams
- sex
- if female, whether or not gravid (with eggs)
- which eye the blood sample was taken from
- percent hematocrit = percent of blood that's red blood cells
- osmolality = a proxy of hydration, should be inversely related to water content of a lizard (this is the average of 1-3 replicates)
- cloacal temperature at the time of CEWL measurement
- processing time for each lizard, when all measurements were finished
- hemolyzed = whether or not red blood cells burst and contaminated plasma

Before loading in this data, some incorrectly-measured hematocrit and osmolality were omitted:
- hematocrit for individuals 1-16, due to observer error
- osmolality for individual 19, due to instrumental error

```{r morpho blood data}
# load and format data
morpho_blood_dat <- read.csv("./data/Herpetology_Data.csv", # filename
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::mutate(# put date and time together
                collect_date_time = (paste(date, collect_time)), 
                # replace some date-time values that have missing times
                collect_date_time = replace(collect_date_time, 
                                            collect_date_time == "4/5/21 NA", NA),
                # correctly format date-time variable
                collect_date_time = as.POSIXct(collect_date_time, 
                                               format = "%m/%d/%y %H:%M"),
                # correctly format date-only variable
                date = as.Date(date, format = "%m/%d/%y"),
                # correctly format collection time variable
                # format extracts just time after posix adds arbitrary date
                collect_time = (as.POSIXct(collect_time, format = "%H:%M")),
                # correctly format processing time variable
                processing_time = (as.POSIXct(processing_time, format = "%H:%M")),
                # set individual_ID variable as a factor, not numeric
                individual_ID = as.factor(individual_ID),
                # set sex variable as a factor, not character
                sex_M_F = as.factor(sex_M_F),
                # set gravidity variable as a factor, not character
                gravid_Y_N = as.factor(gravid_Y_N),
                # set blood sample eye variable as a factor, not character
                blood_sample_eye = as.factor(blood_sample_eye),
                # set hemolyzed variable as a factor, not character
                hemolyzed = as.factor(hemolyzed),
                # compute holding time as capture time - cloacal measurement time:
                hold_time = as.numeric(processing_time - collect_time)
                ) %>%
  # remove two columns not relevant for statistics
  dplyr::select(-sock_ID, -notes)
summary(morpho_blood_dat)
unique(morpho_blood_dat$date)

# export
write.csv(morpho_blood_dat, "exported_data/capture_hydration.csv")
```

I want to test if any IDs are missing, and which ones if so.

```{r}
test <- c(seq(1, 150, by = 1))
lost <- test[test %nin% morpho_blood_dat$individual_ID]
lost
```

Individuals 23 and 56 actually both do not exist because those numbers were skipped when assigning IDs, so we have all the individuals measured in the dataframe.




### CEWL Data

First, load it all in and merge.

Variables in this dataframe are:
- date
- time
- date_time combined variable
- individual_ID for each lizard measured
- region = where on the body CEWL was measured
- TEWL_g_m2h = CEWL measurement value in grams/sq-meter/hour
- ambient_temp_C = temperature when and where measurement was taken
- ambient_RH_percent = relative humidity when and where measurement was taken
- abs_humidity = computed from RH using formula on this website: https://carnotcycle.wordpress.com/2012/08/04/how-to-convert-relative-humidity-to-absolute-humidity/

```{r CEWL data}
# week 1
CEWL_April_05 <- read.csv("./data/capture_CEWL/4-5-21-CEWL.csv", # filename
                          na.strings=c("","NA")) %>% # fix empty cells
  # rename and select the pertinent variables/cols
  # I have to do this for each one 
  # so they all have the same number of columns for joining
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH....
                )

# week 2
CEWL_April_19 <- read.csv("./data/capture_CEWL/4-19-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH....
                )

# week 3
CEWL_April_26 <- read.csv("./data/capture_CEWL/4-26-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH....
                )

# week 4
CEWL_May_3 <- read.csv("./data/capture_CEWL/5-3-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH....
                )

# week 5
CEWL_May_10 <- read.csv("./data/capture_CEWL/5-10-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH....
                )

# week 6
CEWL_May_17 <- read.csv("./data/capture_CEWL/5-17-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Time, Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.., # rename
                ambient_temp_C = AmbT..C., # rename
                ambient_RH_percent = AmbRH....
                )

# merge all CEWL datafiles & reformat
CEWL <- CEWL_April_05 %>% # week 1
  # join with weeks 2-6
  rbind(., CEWL_April_19, 
        CEWL_April_26,
        CEWL_May_3,
        CEWL_May_10,
        CEWL_May_17
        ) %>%
  # remove any unsuccessful measurements
  dplyr::filter(Status == "Normal") %>%
  # extract individual_ID and region separately from the "ID" variable
  separate(ID, c("individual_ID", "region")) %>%
  # reformat data
  dplyr::mutate(# paste and format date-time variable
                CEWL_date_time = as.POSIXct(paste(date, Time), 
                                            format = "%m/%d/%y %I:%M:%S %p"),
                # reformat date only
                date = as.Date(date, format = "%m/%d/%y"),
                # reformat time 
                # format extracts just time after posix adds arbitrary date
                # but then it's a character again... 
                Time = format(as.POSIXct(Time, format = "%I:%M:%S %p"), 
                              format = "%H:%M:%S"),
                # format individual ID as a factor
                individual_ID = as.factor(individual_ID),
                # set body region as a factor variable after getting only the consistent characters due to typos
                region = as.factor(substring(region, 1, 4)),
                # convert RH to absolute humidity
                abs_humidity_g_m3 = ((6.112 * exp((17.67*ambient_temp_C)/(ambient_temp_C + 243.5)) * ambient_RH_percent * 2.1674) / (273.15 + ambient_temp_C))
                ) %>%
  # remove cols not relecant to stats
  dplyr::select(-Status) %>%
  # remove any rows with missing values
  dplyr::filter(complete.cases(.))
summary(CEWL)
```


Write CEWL dataframe as a csv for use in other analyses:

```{r}
write.csv(CEWL, "exported_data/capture_CEWL.csv")
```


Next, split CEWL data by region and compute the average among them.

```{r}
# select each CEWL region separately 
CEWL_dorsum <- CEWL %>%
  dplyr::filter(region == "dors") %>%
  dplyr::select(date, individual_ID, 
                dorsum_TEWL_g_m2h = TEWL_g_m2h)
CEWL_ventrum <- CEWL %>%
  dplyr::filter(region == "vent") %>%
  dplyr::select(date, individual_ID, 
                ventrum_TEWL_g_m2h = TEWL_g_m2h)
CEWL_dewlap <- CEWL %>%
  dplyr::filter(region == "dewl") %>%
  dplyr::select(date, individual_ID, 
                dewlap_TEWL_g_m2h = TEWL_g_m2h)
CEWL_head <- CEWL %>%
  dplyr::filter(region == "head") %>%
  dplyr::select(date, individual_ID, 
                head_TEWL_g_m2h = TEWL_g_m2h)
CEWL_mitepatch <- CEWL %>%
  dplyr::filter(region == "mite") %>%
  dplyr::select(date, individual_ID, 
                mitepatch_TEWL_g_m2h = TEWL_g_m2h)

# also get average across body regions
CEWL_avg <- CEWL %>%
  group_by(individual_ID) %>%
  summarise(avg_CEWL = mean(TEWL_g_m2h)) %>%
  dplyr::select(individual_ID, avg_CEWL)
```


Finally, rearrange and re-join CEWL data.

```{r join data}
# join all CEWL regions to morpho and blood data
data_full <- morpho_blood_dat %>%
  left_join(CEWL_dorsum, by = c("date", "individual_ID")) %>%
  left_join(CEWL_ventrum, by = c("date", "individual_ID")) %>%
  left_join(CEWL_dewlap, by = c("date", "individual_ID")) %>%
  left_join(CEWL_head, by = c("date", "individual_ID")) %>%
  left_join(CEWL_mitepatch, by = c("date", "individual_ID")) %>%
  left_join(CEWL_avg, by = "individual_ID")
```


### Weather Data

This data was obtained from http://www.itrc.org/databases/precip/ (Adcon Server Data) to test the effect of ambient conditions on CEWL. This is different from the ambient conditions already measured with CEWL, which are the temperature and humidity around the measurement device at the time of measurement. We think that the temperature, humidity, wind speed, and solar radiation the lizard was exposed to prior to capture may also affect CEWL.

tbd = daylight savings

```{r}
# load in csvs and put all in one dataframe
weather <- read.csv("./data/weather/4_5Weather.csv", sep = ';') %>%
  rbind(read.csv("./data/weather/4_19Weather.csv", sep = ';')) %>%
  rbind(read.csv("./data/weather/5_3Weather.csv", sep = ';')) %>%
  rbind(read.csv("./data/weather/5_10Weather.csv", sep = ';')) %>%
  rbind(read.csv("./data/weather/5_17Weather.csv", sep = ';')) %>%
  # add a variable for combined date-time
  mutate(collect_date_time = as.POSIXct(paste(Date, Time),
                                format = "%m/%d/%y %I:%M:%S %p")) %>%
  # remove lonely date and time
  dplyr::select(-Date, -Time)
```

The weather data is only every 15 minutes, but I want to match it to any minute measurement, so I need to interpolate the values for each minute.

First, make a separate dataframe with every minute for each of those days.

```{r}
all_times <- data.frame(collect_date_time = c(# April 5
                           seq(from = as.POSIXct("2021-04-05 10:00"),
                               to = as.POSIXct("2021-04-05 16:00"),
                               by="min"),
                           # April 19
                           seq(from = as.POSIXct("2021-04-19 10:00"),
                               to = as.POSIXct("2021-04-19 16:00"),
                               by="min"),
                           # April 26
                           seq(from = as.POSIXct("2021-04-26 10:00"),
                               to = as.POSIXct("2021-04-26 16:00"),
                               by="min"),
                           # May 3
                           seq(from = as.POSIXct("2021-05-03 10:00"),
                               to = as.POSIXct("2021-05-03 16:00"),
                               by="min"),
                           # May 10
                           seq(from = as.POSIXct("2021-05-10 10:00"),
                               to = as.POSIXct("2021-05-10 16:00"),
                               by="min"),
                           # May 17
                           seq(from = as.POSIXct("2021-05-17 10:00"),
                               to = as.POSIXct("2021-05-17 16:00"),
                               by="min")
                           ))

```

Next, merge the weather data into the times dataframe and interpolate the temperature and humidity between measurements.

```{r}
all_times_weather <- all_times %>% # time only dataframe
  # add weather measurements based on matching date-time
  left_join(weather, by = 'collect_date_time') %>%
  # convert temperature units, thanks America
  mutate(temp_C = fahrenheit.to.celsius(Temperature_F, round = 2),
         # interpolate temperatures
         temp_C_interpol = na.approx(temp_C),
         # interpolate humidities
         RH_percent_interpol = na.approx(RH_percent),
         # interpolate Wind Speeds
         Wind_mph_interpol = na.approx(Wind_Speed_mph),
         # interpolate solar radiation
         Solar_rad_Wm2_interpol = na.approx(Pyranometer_W_m),
         # compute absolute humidity
         abs_humidity_g_m3_interpol = ((6.112 * exp((17.67*temp_C_interpol)/(temp_C_interpol + 243.5)) * RH_percent_interpol * 2.1674) / (273.15 + temp_C_interpol))
         ) %>%
  # keep only the relevant variables
  dplyr::select(collect_date_time, 
                temp_C_interpol, 
                RH_percent_interpol, 
                abs_humidity_g_m3_interpol,
                Wind_mph_interpol, 
                Solar_rad_Wm2_interpol)
summary(all_times_weather)
```



### Rain & Humidity

Load data:

```{r}
rain_humd <- read.csv("./data/weather/rain_humidity.csv", sep = ';') %>%
  # add a variable for combined date-time
  mutate(date_time = as.POSIXct(paste(Date, Time),
                                format = "%m/%d/%y %I:%M:%S %p"),
         # fix date only variable format
         Date = as.POSIXct(Date, format = "%m/%d/%y"),
         # convert temperature units, thanks America
         temp_C = fahrenheit.to.celsius(Temp_F, round = 2),
         # compute absolute humidity
         abs_humidity_g_m3 = ((6.112 * exp((17.67*temp_C)/(temp_C + 243.5)) * RH_percent * 2.1674) / (273.15 + temp_C))
         )
summary(rain_humd)
```

Compute cumulative values in the days leading up to lizard capture days:

```{r}
# for April 5
cumul_water_4_05 <- rain_humd %>%
  dplyr::filter(Date < '2021-04-05' & Date > '2021-03-28') %>%
  summarise(total_precip = sum(Precip_inches),
            avg_abs_humd = mean(abs_humidity_g_m3)
            ) %>%
  mutate(sample_date = as.Date("2021-04-05", format = "%Y-%m-%d"))

# for April 19
cumul_water_4_19 <- rain_humd %>%
  dplyr::filter(Date < '2021-04-19' & Date > '2021-04-11') %>%
  summarise(total_precip = sum(Precip_inches),
            avg_abs_humd = mean(abs_humidity_g_m3)
            ) %>%
  mutate(sample_date = as.Date("2021-04-19", format = "%Y-%m-%d"))

# for April 26
cumul_water_4_26 <- rain_humd %>%
  dplyr::filter(Date < '2021-04-26' & Date > '2021-03-18') %>%
  summarise(total_precip = sum(Precip_inches),
            avg_abs_humd = mean(abs_humidity_g_m3)
            ) %>%
  mutate(sample_date = as.Date("2021-04-26", format = "%Y-%m-%d"))

# for May 3
cumul_water_5_03 <- rain_humd %>%
  dplyr::filter(Date < '2021-05-03' & Date > '2021-04-25') %>%
  summarise(total_precip = sum(Precip_inches),
            avg_abs_humd = mean(abs_humidity_g_m3)
            ) %>%
  mutate(sample_date = as.Date("2021-05-03", format = "%Y-%m-%d"))

# for May 10
cumul_water_5_10 <- rain_humd %>%
  dplyr::filter(Date < '2021-05-10' & Date > '2021-05-02') %>%
  summarise(total_precip = sum(Precip_inches),
            avg_abs_humd = mean(abs_humidity_g_m3)
            ) %>%
  mutate(sample_date = as.Date("2021-05-10", format = "%Y-%m-%d"))

# for May 17
cumul_water_5_17 <- rain_humd %>%
  dplyr::filter(Date < '2021-05-17' & Date > '2021-05-09') %>%
  summarise(total_precip = sum(Precip_inches),
            avg_abs_humd = mean(abs_humidity_g_m3)
            ) %>%
  mutate(sample_date = as.Date("2021-05-17", format = "%Y-%m-%d"))

# join them
cumul_water <- cumul_water_4_05 %>%
  rbind(cumul_water_4_19) %>%
  rbind(cumul_water_4_26) %>%
  rbind(cumul_water_5_03) %>%
  rbind(cumul_water_5_10) %>%
  rbind(cumul_water_5_17) %>%
  mutate(prior_rain_Y_N = c("N", "N", "N", "Y", "N", "Y"))
```




### Join Data 

There are several CEWL measurements for each of the other measures, so I'm going to join two ways. Each way allows slightly different analyses.

First, with CEWL as the primary dataframe. This means each of the other variables will be duplicated for each lizards CEWL measurements.

```{r}
all_data_long <- CEWL %>%
  left_join(morpho_blood_dat, 
            by = c("date", "individual_ID")
            ) %>%
  left_join(all_times_weather, 
            by = c("collect_date_time")
            )
```

Second, with morpho_blood_dat as the primary dataframe, and each region's CEWL measurement as an individual column.

```{r}
all_data_wide <- morpho_blood_dat %>%
  left_join(cumul_water, 
            by = c("date" = "sample_date")
            ) %>%
  left_join(all_times_weather, 
            by = c("collect_date_time")
            ) %>%
  left_join(CEWL_dewlap,
            by = c("date", "individual_ID")
            ) %>%
  left_join(CEWL_dorsum,
            by = c("date", "individual_ID")
            ) %>%
  left_join(CEWL_head,
            by = c("date", "individual_ID")
            ) %>%
  left_join(CEWL_mitepatch,
            by = c("date", "individual_ID")
            ) %>%
  left_join(CEWL_ventrum,
            by = c("date", "individual_ID")
            )
```




### Export Data

I want to save the data as csvs for loading into other analyses.

```{r}
write.csv(all_data_long, "exported_data/capture_data_long.csv")
write.csv(all_data_wide, "exported_data/capture_data_wide.csv")
```




# Check Data Distributions

## Histograms & Q-Q Plots

### SVL

not normally distributed, skewed left
 
```{r}
all_data_wide %>%
  ggplot(., aes(x = SVL_mm)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("SVL (mm)") + 
  ylab("Count")
simple.eda(all_data_wide$SVL_mm)
# Normality test if p > .05, data is normal. Data is not normal.
shapiro.test(all_data_wide$SVL_mm)
```

### Mass

slightly skewed left, but nearly a bell curve

```{r}
all_data_wide %>%
  ggplot(., aes(x = mass_g)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Mass (g)") + 
  ylab("Count")
simple.eda(all_data_wide$mass_g)
# Normality test if p > .05, data is normal. Data is not normal.
shapiro.test(all_data_wide$mass_g)
```

### Hematocrit 

looks pretty normally distributed around ~35%, but not statistically normal

```{r}
all_data_wide %>%
  ggplot(., aes(x = hematocrit_percent)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Hematocrit (%)") + 
  ylab("Count")
simple.eda(all_data_wide$hematocrit_percent)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$hematocrit_percent)
```

### Osmolality 

pretty normally distributed around ~370:
only variable to pass normality test (LOL)

```{r}
all_data_wide %>%
  ggplot(., aes(x = osmolality_mmol_kg)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Osmolality (mmol/kg)") + 
  ylab("Count")
simple.eda(all_data_wide$osmolality_mmol_kg)
# Normality test if p > .05, data is normal. Data is normal
shapiro.test(all_data_wide$osmolality_mmol_kg)
```

### Cloacal Temperature 

seems normally distributed, but not normal

```{r}
all_data_wide %>%
  ggplot(., aes(x = cloacal_temp_C)) +
  geom_histogram(color = "black", fill="steelblue", bins=10) + 
  theme_classic() +
  xlab("cloacal temperature (C)") + 
  ylab("Count")
simple.eda(all_data_wide$cloacal_temp_C)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$cloacal_temp_C)
```




### CEWL

```{r}
all_data_long %>%
  ggplot(., aes(x = TEWL_g_m2h)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("TEWL_g_m2h") + 
  ylab("Count")
simple.eda(all_data_long$TEWL_g_m2h)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_long$TEWL_g_m2h)

# Log transformation
shapiro.test(log(all_data_long$TEWL_g_m2h))
# p-value improves to  0.00548, but is still significant
```


### Dewlap CEWL 

very skewed to the right

```{r}
all_data_wide %>%
  ggplot(., aes(x = dewlap_TEWL_g_m2h)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Dewlap CEWL (g/m^2/h)") + 
  ylab("Count")
simple.eda(all_data_wide$dewlap_TEWL_g_m2h)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$dewlap_TEWL_g_m2h)

# Log transformation
shapiro.test(log(all_data_wide$dewlap_TEWL_g_m2h))
# p-value improves to 0.007235, but is still significant

```

### Dorsum CEWL 

slightly skewed to the right

```{r}
all_data_wide %>%
  ggplot(., aes(x = dorsum_TEWL_g_m2h)) +
  geom_histogram(color = "black", fill="steelblue", bins=40) + 
  theme_classic() +
  xlab("Dorsum CEWL (g/m^2/h)") + 
  ylab("Count")
simple.eda(all_data_wide$dorsum_TEWL_g_m2h)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$dorsum_TEWL_g_m2h)

# log transform
shapiro.test(log(all_data_wide$dorsum_TEWL_g_m2h))

```

### Head CEWL 

very skewed to the right

```{r}
all_data_wide %>%
  ggplot(., aes(x = head_TEWL_g_m2h)) +
  geom_histogram(color = "black", fill="steelblue", bins=40) + 
  theme_classic() +
  xlab("Head CEWL (g/m^2/h)") + 
  ylab("Count")
simple.eda(all_data_wide$head_TEWL_g_m2h)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$head_TEWL_g_m2h)

# Log transform made data normal
shapiro.test(log(all_data_wide$head_TEWL_g_m2h))

```

### Mite Patch CEWL 

very skewed to the right

```{r}
all_data_wide %>%
  ggplot(., aes(x = mitepatch_TEWL_g_m2h)) +
  geom_histogram(color = "black", fill="steelblue", bins=40) + 
  theme_classic() +
  xlab("Mite Patch CEWL (g/m^2/h)") + 
  ylab("Count")
simple.eda(all_data_wide$mitepatch_TEWL_g_m2h)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$mitepatch_TEWL_g_m2h)

# Log transform p-0.04107, close to being normal
shapiro.test(log(all_data_wide$mitepatch_TEWL_g_m2h))
```

### Ventrum CEWL 

slightly skewed to the right, somewhat evenly distributed, but not statistically normal

```{r}
all_data_wide %>%
  ggplot(., aes(x = ventrum_TEWL_g_m2h)) +
  geom_histogram(color = "black", fill="steelblue", bins=30) + 
  theme_classic() +
  xlab("Ventrum CEWL (g/m^2/h)") + 
  ylab("Count")
simple.eda(all_data_wide$ventrum_TEWL_g_m2h)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$ventrum_TEWL_g_m2h)

# log transform didnt work 0.0006058
shapiro.test(log(all_data_wide$ventrum_TEWL_g_m2h))
```

### Temperature 

not normal

```{r}
all_data_wide %>%
  ggplot(., aes(x = temp_C_interpol)) +
  geom_histogram(color = "black", fill="steelblue", bins=30) + 
  theme_classic() +
  xlab("Temperature (C)") + 
  ylab("Count")
simple.eda(all_data_wide$temp_C_interpol)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$temp_C_interpol)

# log transform 
shapiro.test(log(all_data_wide$temp_C_interpol))
```

### Humidity

not normal

```{r}
all_data_wide %>%
  ggplot(., aes(x = abs_humidity_g_m3_interpol)) +
  geom_histogram(color = "black", fill="steelblue", bins=30) + 
  theme_classic() +
  xlab("Absolute Humidity (g/m^3)") + 
  ylab("Count")
simple.eda(all_data_wide$abs_humidity_g_m3_interpol)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$abs_humidity_g_m3_interpol)

# Doesn't fix non normality
shapiro.test(log(all_data_wide$abs_humidity_g_m3_interpol))

```

### Wind

not normal

```{r}
all_data_wide %>%
  ggplot(., aes(x = Wind_mph_interpol)) +
  geom_histogram(color = "black", fill="steelblue", bins=30) + 
  theme_classic() +
  xlab("Wind Speed (mph)") + 
  ylab("Count")
simple.eda(all_data_wide$Wind_mph_interpol)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$Wind_mph_interpol)

# ln transformation doesn't fix non normality
shapiro.test(log(all_data_wide$Wind_mph_interpol))
```

### Solar Radiation

not normal

```{r}
all_data_wide %>%
  ggplot(., aes(x = Solar_rad_Wm2_interpol)) +
  geom_histogram(color = "black", fill="steelblue", bins=30) + 
  theme_classic() +
  xlab("Solar Radiation (W/m^2)") + 
  ylab("Count")
simple.eda(all_data_wide$Solar_rad_Wm2_interpol)
# Normality test if p > .05, data is normal. Data is not normal
shapiro.test(all_data_wide$Solar_rad_Wm2_interpol)

# Doesn't fix non normality
shapiro.test(log(all_data_wide$Solar_rad_Wm2_interpol))
```

## Conclusion

Osmolality was the only normally distributed variable. 

The following variables each had non-normal distributions:
- SVL (skewed left)
- mass (skewed left)
- hct (skewed both ways??)
- cloacal temp (skewed right)
- CEWL overall (skewed right)
- CEWL for each region individually (skewed right)
- capture temp (multimodal)
- capture humidity (multimodal and skewed left)
- wind speed (multimodal)
- solar radiation (multimodal)

This will be important to keep in mind going forward. Once we have our model, we will need to very very carefully check assumptions.



# Basic Figs & GLMs

## What affects hydration & health?

Potential relationships:
- Hct or Osml ~ SVL, mass, sex, gravidity, eye, hemolyzed, week

### Hct ~ SVL

- No sig relationship

```{r}
# plot
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = SVL_mm,
                 y = hematocrit_percent, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = SVL_mm, 
                  y = hematocrit_percent, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("SVL") + 
  ylab("Hct") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_hct_SVL <- lm(hematocrit_percent ~ SVL_mm,
           data = all_data_wide)
summary(glm_hct_SVL)
```



### Osml ~ SVL

- no sig relationship

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = SVL_mm,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = SVL_mm, 
                  y = osmolality_mmol_kg, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("SVL") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_osml_SVL <- lm(osmolality_mmol_kg ~ SVL_mm,
           data = all_data_wide)
summary(glm_osml_SVL)
```


### Hct ~ Mass

- no sig relationship

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = mass_g,
                 y = hematocrit_percent, 
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = mass_g, 
                  y = hematocrit_percent, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Mass") + 
  ylab("Hematocrit") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_hct_mass <- lm(hematocrit_percent ~ mass_g,
           data = all_data_wide)
summary(glm_hct_mass)
```

### Osml ~ Mass

- no sig relationship

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = mass_g,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = mass_g, 
                  y = osmolality_mmol_kg), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "gray",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Mass") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_osml_mass <- lm(osmolality_mmol_kg ~ mass_g,
           data = all_data_wide)
summary(glm_osml_mass)
```


### Hct ~ Sex

- sig relationship, males have a higher hematocrit %

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = sex_M_F, 
                   y = hematocrit_percent, 
                   color = sex_M_F
                   ), 
               size = 1,
               alpha = 1) + 
  theme_classic() + 
  xlab("Sex") + 
  ylab("Hematocrit (%)") + 
  annotate("text", x = 1.5, y = 45, 
           label = "paste(italic(p), \" = 0.0152\")", 
           parse = TRUE,
           size = 6) +
  ylim(10, 50) +
  scale_x_discrete(labels = c("F" = "Female",
                              "M" = "Male")) +
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
) -> hct_sex_fig
hct_sex_fig

# export figure
#ggsave(filename = "hct_sex_fig.tiff",
 #      plot = hct_sex_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)

# glms
aov_hct_sex <- aov(hematocrit_percent ~ sex_M_F,
           data = all_data_wide)
TukeyHSD(aov_hct_sex) # this is the stat to present
glm_hct_sex_mass <- glm(hematocrit_percent ~ sex_M_F*mass_g - mass_g,
           data = all_data_wide)
summary(glm_hct_sex_mass)
```

Hematocrit is significantly predicted by sex, but the interaction between sex and mass is ~nonexistent. 


### Osml ~ Sex

females have higher osmolarity (nonsig)

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = sex_M_F, 
                   y = osmolality_mmol_kg, 
                   color = sex_M_F
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Sex", 
                      values = c("green4", "salmon1") ) +
  theme_classic() + 
  xlab("Sex") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_osml_sex <- lm(osmolality_mmol_kg ~ sex_M_F,
           data = all_data_wide)
summary(glm_osml_sex)
```

### Hct ~ Gravidity

- no sig diff

```{r}
all_data_wide %>% 
  dplyr::filter(sex_M_F == 'F') %>%
  ggplot(data = .) + 
  geom_boxplot(aes(x = gravid_Y_N, 
                   y = hematocrit_percent, 
                   color = gravid_Y_N
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Gravid", 
                      values = c("green4", "salmon1") ) +
  theme_classic() + 
  xlab("Gravid (Females Only)") + 
  ylab("Hematocrit") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_hct_gravid <- lm(hematocrit_percent ~ gravid_Y_N,
           data = all_data_wide)
summary(glm_hct_gravid)
```

### Osml ~ Gravidity

gravid F have higher osmolarity (nonsig)

```{r}
all_data_wide %>% 
  dplyr::filter(sex_M_F == 'F') %>%
  ggplot(data = .) + 
  geom_boxplot(aes(x = gravid_Y_N, 
                   y = osmolality_mmol_kg, 
                   color = gravid_Y_N
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Gravid", 
                      values = c("green4", "salmon1") ) +
  theme_classic() + 
  xlab("Gravid (Females Only)") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_osml_gravid <- lm(osmolality_mmol_kg ~ gravid_Y_N,
           data = all_data_wide)
summary(glm_osml_gravid)
```

### Hct ~ Sample Eye

Actually, any blood samples not taken from the right eye ended up being excluded or didn't have hematocrit values, so we can't test this difference.

```{r}
all_data_wide %>% 
  dplyr::filter(blood_sample_eye %in% c("R", "L")) %>%
  ggplot(data = .) + 
  geom_boxplot(aes(x = blood_sample_eye, 
                   y = hematocrit_percent, 
                   color = blood_sample_eye
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Blood Sample Eye", 
                      values = c("green4", "salmon1") ) +
  theme_classic() + 
  xlab("Blood Sample Eye") + 
  ylab("Hematocrit") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# can't do a glm
```


### Osml ~ Sample Eye

```{r}
all_data_wide %>% 
  dplyr::filter(blood_sample_eye %in% c("R", "L")) %>%
  ggplot(data = .) + 
  geom_boxplot(aes(x = blood_sample_eye, 
                   y = osmolality_mmol_kg, 
                   color = blood_sample_eye
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Blood Sample Eye", 
                      values = c("green4", "salmon1") ) +
  theme_classic() + 
  xlab("Blood Sample Eye") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_osml_eye <- lm(osmolality_mmol_kg ~ blood_sample_eye,
           data = all_data_wide)
summary(glm_osml_eye)
```

I wasn't expecting this, and I'm not sure whether to actually include it or not. 

### Hct ~ Hemolyzed/Not

- no sig diff

```{r}
all_data_wide %>% 
  dplyr::filter(hemolyzed %in% c("Y", "N")) %>%
  ggplot(data = .) + 
  geom_boxplot(aes(x = hemolyzed, 
                   y = hematocrit_percent, 
                   color = hemolyzed
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Blood Sample Eye", 
                      values = c("green4", "salmon1", "green4", "salmon1") ) +
  theme_classic() + 
  xlab("Whether or not Sample Hemolyzed") + 
  ylab("Hematocrit") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_hct_hem <- lm(hematocrit_percent ~ hemolyzed,
           data = all_data_wide)
summary(glm_hct_hem)
```



### Osml ~ Hemolyzed/Not

```{r}
all_data_wide %>% 
  dplyr::filter(hemolyzed %in% c("Y", "N")) %>%
  ggplot(data = .) + 
  geom_boxplot(aes(x = hemolyzed, 
                   y = osmolality_mmol_kg, 
                   color = hemolyzed
                   ), 
               size = 1,
               alpha = 0.6) + 
  scale_colour_manual(name = "Blood Sample Eye", 
                      values = c("green4", "salmon1", "green4", "salmon1") ) +
  theme_classic() + 
  xlab("Whether or not Sample Hemolyzed") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_osml_hem <- lm(osmolality_mmol_kg ~ hemolyzed,
           data = all_data_wide)
summary(glm_osml_hem)
```



### Hct ~ Week

- no sig diff

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = as.factor(date), 
                   y = hematocrit_percent, 
                   color = as.factor(date)
                   ), 
               size = 1,
               alpha = 0.6) + 
  theme_classic() + 
  xlab("Sampling Date") + 
  ylab("Hematocrit") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_hct_wk <- lm(hematocrit_percent ~ date,
           data = all_data_wide)
summary(glm_hct_wk)
```



### Osml ~ Week

- Sig difference in osmolarity by week

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = as.factor(date), 
                   y = osmolality_mmol_kg, 
                   color = as.factor(date)
                   ), 
               size = 1,
               alpha = 0.6) + 
  theme_classic() + 
  xlab("Date") + 
  ylab("Osmolality (mmol / kg)") + 
  annotate("text", x = 1, y = 365, label = "a", size = 6) +
  annotate("text", x = 2, y = 390, label = "b", size = 6) +
  annotate("text", x = 3, y = 437, label = "c", size = 6) +
  annotate("text", x = 4, y = 427, label = "c", size = 6) +
  annotate("text", x = 5, y = 417, label = "d", size = 6) +
  annotate("text", x = 6, y = 447, label = "d", size = 6) +
  scale_x_discrete(labels = c("2021-04-05" = "Apr 5", 
                              "2021-04-19" = "Apr 19",
                              "2021-04-26" = "Apr 26",
                              "2021-05-03" = "May 3", 
                              "2021-05-10" = "May 10",
                              "2021-05-17" = "May 17")) +
  scale_color_brewer(palette = "Set2") +
  ylim(300, 450) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
) -> osml_date_fig
osml_date_fig

# export figure
#ggsave(filename = "osml_date_fig.tiff",
 #      plot = osml_date_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)

# glm
glm_osml_wk <- glm(osmolality_mmol_kg ~ date,
           data = all_data_wide)
summary(glm_osml_wk)
```

There was a little rain on April 27.... could that be why there was a slight drop in osmolality?! Not sure if the May 3-> 10 drop was because of rain or fixing the osmometer, though... Need to look for better daily rainfall records, maybe look in relation to humidity?.

distinguish pairwise differences using an ANOVA:

```{r}
osml_date_aov <- aov(osmolality_mmol_kg ~ as.factor(date), 
                     data = all_data_wide)
TukeyHSD(osml_date_aov)
```


### Humidity ~ Week

- Significant difference in humidity by week

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = as.factor(date), 
                   y = RH_percent_interpol, 
                   color = as.factor(date)
                   ), 
               size = 1,
               alpha = 0.6) + 
  theme_classic() + 
  xlab("Sampling Date") + 
  ylab("Relative Humidity") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_hum_wk <- glm(RH_percent_interpol ~ date,
           data = all_data_wide)
summary(glm_hum_wk)

```

try for absolute humidity:

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = as.factor(date), 
                   y = abs_humidity_g_m3_interpol, 
                   color = as.factor(date)
                   ), 
               size = 1,
               alpha = 0.6) + 
  theme_classic() + 
  xlab("Sampling Date") + 
  ylab("Absolute Humidity") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm_abshum_wk <- glm(abs_humidity_g_m3_interpol ~ as.factor(date),
           data = all_data_wide)
summary(glm_abshum_wk)
aov_abshum_wk <- aov(abs_humidity_g_m3_interpol ~ as.factor(date),
           data = all_data_wide)
TukeyHSD(aov_abshum_wk)
```


### Osml, Humidity, Week

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = as.factor(date), 
                   y = osmolality_mmol_kg, 
                   color = as.factor(date)
                   ), 
               size = 1,
               alpha = 0.6) + 
  geom_jitter(aes(x = as.factor(date), 
                   y = 35*abs_humidity_g_m3_interpol)) +
  theme_classic() + 
  xlab("Sampling Date") + 
  ylab("Osmolality (mmol/kg)") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)
```



### Osml ~ R. Humidity

- Sig difference, negative relationship. With every percent increase in relative humidity osmolarity drops 1.2.

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = RH_percent_interpol,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = RH_percent_interpol, 
                  y = osmolality_mmol_kg), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "gray",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Relative Humidity") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_osml_hum <- lm(osmolality_mmol_kg ~ RH_percent_interpol,
           data = all_data_wide)
summary(glm_osml_hum)
```

That's a pretty good relationship... 


### Osml ~ Abs. Humidity

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = abs_humidity_g_m3_interpol,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.4) + 
  stat_smooth(aes(x = abs_humidity_g_m3_interpol, 
                  y = osmolality_mmol_kg), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "royalblue",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab(bquote('Absolute Humidity at Capture (g / '*m^2*')')) + 
  ylab("Osmolality (mmol / kg)") + 
  xlim(8, 12) +
  ylim(300, 450) +
  annotate("text", x = 8.5, y = 440, 
           label = "paste(italic(R) ^ 2, \" = 0.036\")", 
           parse = TRUE,
           size = 6) +
  annotate("text", x = 8.5, y = 428, 
           label = "paste(italic(p), \" = 0.0126\")", 
           parse = TRUE,
           size = 6) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
) -> osml_abhum_fig
osml_abhum_fig

# export figure
#ggsave(filename = "osml_abhum_fig.tiff",
 #      plot = osml_abhum_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)

# glm
glm_osml_abshum <- lm(osmolality_mmol_kg ~ abs_humidity_g_m3_interpol,
           data = all_data_wide)
summary(glm_osml_abshum)
```

*positive* correlation...


### Osml ~ Avg. Abs. Humidity

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = avg_abs_humd,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = avg_abs_humd, 
                  y = osmolality_mmol_kg), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "gray",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Prior Week's Average Absolute Humidity") + 
  ylab("Osmolality (mmol/kg)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_osml_avgabshum <- lm(osmolality_mmol_kg ~ avg_abs_humd,
           data = all_data_wide)
summary(glm_osml_avgabshum)
```

*positive* correlation...



### Osml ~ Rain

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = total_precip,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = total_precip, 
                  y = osmolality_mmol_kg), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "gray",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Prior Week's Total Precipitation (inches)") + 
  ylab("Osmolality (mmol/kg)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)
```

This isn't really helpful considering how little variation in precipitation there was.

Maybe try a binary predictor of whether or not there was rain in the week prior to sampling:

```{r}
# ANOVA
osml_rain_aov <- aov(data = all_data_wide, 
                     osmolality_mmol_kg ~ prior_rain_Y_N)
TukeyHSD(osml_rain_aov)

# plot
all_data_wide %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = prior_rain_Y_N, 
                   y = osmolality_mmol_kg, 
                   color = prior_rain_Y_N
                   ), 
               size = 1,
               alpha = 0.6) + 
  theme_classic() + 
  xlab("Whether it Rained in the Last Week") + 
  ylab("Osmolality (mmol/kg)") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)
```

Wow, very counter to our predictions.






### Hct ~ Humidity

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = RH_percent_interpol,
                 y = hematocrit_percent, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = RH_percent_interpol, 
                  y = hematocrit_percent), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "blue",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Ambient Relative Humidity at Capture (%)") + 
  ylab("Hematocrit (%)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_hct_RH <- lm(hematocrit_percent ~ RH_percent_interpol,
           data = all_data_wide)
summary(glm_hct_RH)
```


### Hct ~ Temperature

as hematocrit increases ambient temp increases (nonsig)

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = temp_C_interpol,
                 y = hematocrit_percent, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = temp_C_interpol, 
                  y = hematocrit_percent), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "maroon",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Ambient Temperature at Capture (C)") + 
  ylab("Hematocrit (%)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_hct_temp <- lm(hematocrit_percent ~ temp_C_interpol,
           data = all_data_wide)
summary(glm_hct_temp)
```

### Osml ~ Temperature

- very strong positive relationship

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = temp_C_interpol,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.4) + 
  stat_smooth(aes(x = temp_C_interpol, 
                  y = osmolality_mmol_kg), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              color = "maroon",
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Ambient Temperature at Capture (C)") + 
  ylab("Osmolality (mmol / kg)") + 
  annotate("text", x = 22, y = 320, 
           label = "paste(italic(R) ^ 2, \" = 0.2952\")", 
           parse = TRUE,
           size = 6) +
  annotate("text", x = 22, y = 308, 
           label = "paste(italic(p), \" < 0.0001\")", 
           parse = TRUE,
           size = 6) +
  xlim(14, 24) +
  ylim(300,450) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
) -> osml_temp_fig
osml_temp_fig

# export figure
#ggsave(filename = "osml_temp_fig.tiff",
 #      plot = osml_temp_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)

# glm
glm_osml_temp <- lm(osmolality_mmol_kg ~ temp_C_interpol,
           data = all_data_wide)
summary(glm_osml_temp)
```

### Hct ~ Individual

- not sig

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = individual_ID,
                 y = hematocrit_percent, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = individual_ID, 
                  y = hematocrit_percent, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Individual Lizard") + 
  ylab("Hematocrit (%)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_hct_ID <- lm(hematocrit_percent ~ individual_ID,
           data = all_data_wide)
summary(glm_hct_ID)
```

### Osml ~ Individual

- sig relationship when numeric variable...

```{r}
all_data_wide %>% 
  ggplot(data = .) + 
  geom_point(aes(x = individual_ID,
                 y = osmolality_mmol_kg, 
                 ), 
             size = 1, 
             alpha = 0.6) + 
  stat_smooth(aes(x = individual_ID, 
                  y = osmolality_mmol_kg, 
                  ), 
              formula = y ~ x, 
              method = "lm", 
              color = "gray",
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Individual Lizard") + 
  ylab("Osmolality") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm_osml_ID <- lm(osmolality_mmol_kg ~ individual_ID,
           data = all_data_wide)
summary(glm_osml_ID)
```



## Conclusion

Hydration seems to be affected by:
- mass (NS)
- sex (NS)
- gravid/not (NS)
- sample eye (*) !
- whether or not the sample was hemolyzed (*) !
- week/date of sampling!! (***)
- individual (*** but likely confounded with week/date...)
- capture temp & RH (both ***)

So, for the LMM to predict osmolality, we will start with sample eye, hemolyzed/not, week/date, individual, and capture temp and absolute humidity as our predictor variables in the model. 

Hematocrit seems to be affected by:
- mass (NS)
- sex (*)
- capture temp & RH (NS)

So we will only include the model for hct ~ sex in the paper. 



## What affects evaporative water loss?

Potential relationships:
- CEWL ~ date/week, 
        individual, 
        SVL, 
        mass, 
        gravidity, 
        hct, 
        osml, 
        cloacal temp, 
        ambient temp, 
        ambient RH, 
        measurement temp,
        measurement RH,
        **body region**

### CEWL ~ Body Region

figure:

```{r}
CEWL %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = reorder(region, TEWL_g_m2h), 
                   y = TEWL_g_m2h, 
                   color = region
                   ), 
               size = 1,
               alpha = 1) +
  scale_x_discrete(labels = c("Dewlap", "Dorsum",
                              "Mite Patch", "Head", "Ventrum")) +
  theme_classic() + 
  xlab("Body Region") + 
  ylab(bquote('CEWL (g / '*m^2~h*')')) + 
  annotate("text", x = 1, y = 65, label = "a", size = 6) +
  annotate("text", x = 2, y = 62, label = "a", size = 6) +
  annotate("text", x = 3, y = 105, label = "b", size = 6) +
  annotate("text", x = 4, y = 92, label = "b", size = 6) +
  annotate("text", x = 5, y = 75, label = "b", size = 6) +
  scale_color_brewer(palette = "Set2") +
  ylim(1, 110) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
) -> CEWL_region_fig
CEWL_region_fig

# export figure
#ggsave(filename = "CEWL_region_fig.tiff",
 #      plot = CEWL_region_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)
```


stats:

```{r}
# GLM
glm1 <- lm(TEWL_g_m2h ~ region, data = CEWL)
summary(glm1)
```


I think that the way the GLM tests differences is not preferable. It tests how different each category is from the one reference category, rather than testing differences of each region against each other region, in a pairwise test. So, I will try an ANOVA to see if I can get those pairwise statistics.

```{r}
# one-way ANOVA
CEWL_region_aov <- aov(data = all_data_long,
                       TEWL_g_m2h ~ region)
# post-hoc pairwise analysis
TukeyHSD(CEWL_region_aov)
```

Okay, the GLM coefficients and ANOVA differences are the same, so either way should be fine, but helpful to have complete pairwise stats from ANOVA for marking the figure.



### CEWL ~ Osmolality

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = osmolality_mmol_kg,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Osmolality") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm2 <- lm(TEWL_g_m2h ~ region + osmolality_mmol_kg,
           data = all_data_long)
summary(glm2)


# Facet ggplot
ggplot(aes(osmolality_mmol_kg, TEWL_g_m2h), data = all_data_long) + 
  geom_point() + 
  stat_smooth(aes(x = osmolality_mmol_kg, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 )+
    theme_classic() + 
  xlab("Osmolality") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)+
  facet_wrap(~ region) # create a facet for each body region
```


### CEWL ~ Hematocrit

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = hematocrit_percent,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = hematocrit_percent, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Hematocrit") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glms
# CEWL ~ region + hct
glm3 <- lm(TEWL_g_m2h ~ region + hematocrit_percent,
           data = all_data_long)
summary(glm3)
```


### CEWL ~ Cloacal Temperature

figure:

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = cloacal_temp_C,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.4) + 
  stat_smooth(aes(x = cloacal_temp_C, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1) + 
  theme_classic() + 
  xlab("Cloacal Temperature (C)") + 
  ylab(bquote('CEWL (g / '*m^2~h*')')) + 
  #annotate("text", x = 1, y = 65, label = "a", size = 6) +
  scale_color_brewer(palette = "Set2",
                     labels = c("Dewlap", "Dorsum", "Head",
                              "Mite Patch", "Ventrum"),
                     name = "") +
  ylim(1, 100) +
  xlim(20, 28) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text.align = 0,
        legend.position = c(0.9, 0.85)
  #) +
  #guides(color = guide_legend(nrow = 2, byrow = TRUE)
         ) -> CEWL_ctemp_fig
CEWL_ctemp_fig

# export figure
#ggsave(filename = "CEWL_ctemp_fig.tiff",
 #      plot = CEWL_ctemp_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)
```

stats:

```{r}
# glms
# CEWL ~ region + ctemp
glm4 <- lm(TEWL_g_m2h ~ region + cloacal_temp_C,
           data = all_data_long)
summary(glm4)
```



### CEWL ~ Capture Temperature

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = temp_C_interpol,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = temp_C_interpol, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Ambient Temperature at Capture") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glms
# CEWL ~ region + capture temp
glm7 <- lm(TEWL_g_m2h ~ region + temp_C_interpol,
           data = all_data_long)
summary(glm7)
```


### CEWL ~ Capture Humidity

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = RH_percent_interpol,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = RH_percent_interpol, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Ambient Relative Humidity at Capture") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glms
# CEWL ~ region + capture RH
glm8 <- lm(TEWL_g_m2h ~ region + RH_percent_interpol,
           data = all_data_long)
summary(glm8)
```


### CEWL ~ Abs Humidity

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_jitter(aes(x = abs_humidity_g_m3_interpol,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.4) + 
  stat_smooth(aes(x = abs_humidity_g_m3_interpol, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab('Absolute Humidity at Capture (g / '*m^3~')') + 
  ylab(bquote('CEWL (g / '*m^2~h*')')) + 
  #annotate("text", x = 1, y = 65, label = "a", size = 6) +
  scale_color_brewer(palette = "Set2",
                     labels = c("Dewlap", "Dorsum",
                              "Head", "Mite Patch", "Ventrum"),
                     name = "") +
  ylim(1, 100) +
  xlim(8, 12) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text.align = 0,
        legend.position = c(0.15, 0.85)
         ) -> CEWL_abshum_fig
CEWL_abshum_fig

# export figure
#ggsave(filename = "CEWL_abshum_fig.tiff",
 #      plot = CEWL_abshum_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)
```

model: 

```{r}
glm8.1 <- lm(TEWL_g_m2h ~ region + abs_humidity_g_m3_interpol,
           data = all_data_long)
summary(glm8.1)
```


### CEWL ~ Measurement Temperature

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = ambient_temp_C,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = ambient_temp_C, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Ambient Temperature During Measurement") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
# CEWL ~ region + aquaflux measurement temp
glm9 <- lm(TEWL_g_m2h ~ region + ambient_temp_C,
           data = all_data_long)
summary(glm9)
```


### CEWL ~ Measurement Humidity

Very interesting relationship! Mite patch CEWL decreases as ambient humidity increases, but every other location appears to increase. In this case, an interaction term is warranted.

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = ambient_RH_percent,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = ambient_RH_percent, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Ambient Relative Humidity During Measurement") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
# CEWL ~ region + aquaflux measurement RH
glm9 <- lm(TEWL_g_m2h ~ region * ambient_RH_percent,
           data = all_data_long)
summary(glm9)
```

### CEWL ~ Wind Speed

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = Wind_mph_interpol,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = Wind_mph_interpol, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Average Windspeed During Measurement") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
# CEWL ~ region + aquaflux measurement RH
glm10 <- lm(TEWL_g_m2h ~ region + Wind_mph_interpol,
           data = all_data_long)
summary(glm10)
```

### CEWL ~ Solar Rad

- Definitely could have an effect may be worth testing in the model

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = Solar_rad_Wm2_interpol,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = Solar_rad_Wm2_interpol, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Solar Radiation W/m^2") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
# CEWL ~ region + aquaflux measurement RH
glm11 <- lm(TEWL_g_m2h ~ region + Solar_rad_Wm2_interpol,
           data = all_data_long)
summary(glm11)
```


### CEWL ~ Individual

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = individual_ID,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = individual_ID, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Individual Lizard") + 
  ylab("CEWL") + 
  
  # just to get a better look
  # ylim(5, 40) +
  
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm6 <- lm(TEWL_g_m2h ~ region + individual_ID,
           data = all_data_long)
summary(glm6)
```

### CEWL ~ SVL

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = SVL_mm,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = SVL_mm, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("SVL") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

# glm
glm10 <- lm(TEWL_g_m2h ~ region + SVL_mm,
           data = all_data_long)
summary(glm10)
```

### CEWL ~ Mass

Head has an opposite trend from all the other body regions, so we need an interaction term.

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = mass_g,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.4) + 
  #scale_colour_manual(values = c("palegreen4", "lightblue4", 
   #                              "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = mass_g, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Mass (g)") + 
  ylab(bquote('CEWL (g / '*m^2~h*')')) + 
  #annotate("text", x = 1, y = 65, label = "a", size = 6) +
  scale_color_brewer(palette = "Set2",
                     labels = c("Dewlap", "Dorsum",
                              "Head", "Mite Patch", "Ventrum"),
                     name = "") +
  ylim(1, 100) +
  xlim(2, 16) +
  scale_x_continuous(breaks = c(seq(2, 16, by = 2))) +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text.align = 0,
        legend.position = c(0.15, 0.85)
         ) -> CEWL_mass_fig
CEWL_mass_fig

# export figure
#ggsave(filename = "CEWL_mass_fig.tiff",
 #      plot = CEWL_mass_fig,
  #     path = "./final_figures",
   #    device = "tiff",
    #   dpi = 1200,
     #  width = 6, height = 4)


# glm
glm11 <- lm(TEWL_g_m2h ~ region*mass_g,
           data = all_data_long)
summary(glm11)
```

### CEWL ~ Sex

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = sex_M_F, 
                   y = TEWL_g_m2h, 
                   color = sex_M_F
                   ), 
               size = 1,
               alpha = 0.6) + 
  facet_wrap(~region) +
  scale_color_manual(values = c("royalblue1", "mediumorchid")) +
  scale_x_discrete(breaks = c(1,2,3)) +
  theme_classic() + 
  xlab("Sex") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm5 <- lm(TEWL_g_m2h ~ region + sex_M_F, 
           data = all_data_long)
summary(glm5)
```

### CEWL ~ Gravidity

```{r}
all_data_long %>% 
  dplyr::filter(sex_M_F == "F") %>%
  ggplot(data = .) + 
  geom_boxplot(aes(x = gravid_Y_N, 
                   y = TEWL_g_m2h, 
                   color = gravid_Y_N
                   ), 
               size = 1,
               alpha = 0.6) + 
  facet_wrap(~region) +
  scale_color_manual(values = c("royalblue1", "mediumorchid")) +
  scale_x_discrete(breaks = c(1,2,3)) +
  theme_classic() + 
  xlab("Female Gravid or Not") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm5 <- lm(TEWL_g_m2h ~ region + gravid_Y_N, 
           data = all_data_long)
summary(glm5)
```


### CEWL ~ Week

```{r}
CEWL %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = as.factor(date), 
                   y = TEWL_g_m2h, 
                   color = as.factor(date)
                   ), 
               size = 1,
               alpha = 0.6) + 
  facet_wrap(~region) + # could not figure out how to change facet labels without changing underlying data
  scale_color_manual(values = c("royalblue1", "mediumorchid", "seagreen4",
                                "royalblue1", "mediumorchid", "seagreen4")) +
  scale_x_discrete(breaks = c(1,2,3)) +
  theme_classic() + 
  xlab("Date") + 
  ylab("CEWL") + 
  theme(text = element_text(color = "black", family = "sans", size = 12),
        axis.text = element_text(color = "black", family = "sans", size = 10),
        legend.text.align = 0,
        legend.position = "none"
)

# glm
glm12 <- lm(TEWL_g_m2h ~ region + date,
           data = all_data_long)
summary(glm12)
```


### CEWL ~ holding time

```{r}
all_data_long %>% 
  ggplot(data = .) + 
  geom_point(aes(x = hold_time,
                 y = TEWL_g_m2h, 
                 color = region
                 ), 
             size = 1, 
             alpha = 0.6) + 
  scale_colour_manual(values = c("palegreen4", "lightblue4", 
                                 "plum4", "lightpink3", "seashell4")) +
  stat_smooth(aes(x = hold_time, 
                  y = TEWL_g_m2h, 
                  color = region
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  xlab("Holding Time (minutes)") + 
  ylab("CEWL (g/m^2/hr)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
)

htime_CEWL_glm <- glm(data = all_data_long, 
                      TEWL_g_m2h ~ hold_time + region)
summary(htime_CEWL_glm)
```

Holding time is still significant, even after regional differences are accounted for. 


## Conclusion

CEWL is/is not affected by:
- *body region* - significant!
- osmolality (hydration) - not significant
- hematocrit (health) - not significant 
- *cloacal temperature* at measurement - sig!
- *capture temp* - sig!
- capture RH/absH - sig! but we will only use *absolute humidity* since it's decoupled from temperature
- measurement temperature and humidity - nonsig
- capture *wind speed* and *solar radiation* - sig!
- individual ID - nonsig
- *mass & SVL* - both sig!
- sex & gravidity - nonsig
- week/date - nonsig as a standalone variable
- *hold time* - significant!


# LMMs

## Hydration

Based on the simple linear models and figures above, osmolality should be predicted by sample eye, hemolysis, date/week, individual, capture temperature, and capture absolute humidity.

Prep dataframe for computing models:

```{r}
hydrat_mod_dat <- all_data_wide %>%
  dplyr::select(date, 
                individual_ID,
                osmolality_mmol_kg,
                temp_C_interpol,
                abs_humidity_g_m3_interpol,
                blood_sample_eye,
                hemolyzed
                ) %>%
  dplyr::filter(complete.cases(.))
```


### Multicollinearity

First, check for multicollinearity among independent variables:

```{r}
pairs(hydrat_mod_dat)

# also make another plot with r-sq values
# only continuous numeric variables work for this one
hydrat_mod_dat %>% 
  # select variables of interest
  dplyr::select(temp_C_interpol,
                abs_humidity_g_m3_interpol) %>% 
  # multicollinearity plot
  chart.Correlation(., histogram = F, pch = 19)
```

Date and individual_ID are collinear and should not both be used as fixed effects. Temperature and absolute humidity are the only continuous variables, and they are not badly collinear.



### Models & Selection
                
Start with all the variables that were significant individually in SLRs. 

```{r, hydration models}
# model 1
hydrat_mod1 <- lme4::lmer(data = hydrat_mod_dat,
                          osmolality_mmol_kg ~ 
                           abs_humidity_g_m3_interpol*temp_C_interpol +
                            date + blood_sample_eye + hemolyzed +
                            (1|individual_ID)) 
summary(hydrat_mod1)
```

I think I should probably skip blood_sample_eye and hemolyzed in the model. I can note they were significant in the paper as SLRs, but I think they might be one of the problems throwing warnings in the model. 

```{r, hydration model 2}
# model 2
hydrat_mod2 <- lme4::lmer(data = hydrat_mod_dat,
                          osmolality_mmol_kg ~ 
                           abs_humidity_g_m3_interpol*temp_C_interpol +
                            date + 
                            (1|individual_ID)) 
summary(hydrat_mod2)
```

Still having some of the same errors... try scaling the continuous variables. 

```{r}
hydrat_mod_dat_scaled <- hydrat_mod_dat %>%
  mutate(osml_scaled = scale(osmolality_mmol_kg),
         temp_scaled = scale(temp_C_interpol),
         abshum_scaled = scale(abs_humidity_g_m3_interpol))
```


```{r, hydration model 3}
# model 3
hydrat_mod3 <- lme4::lmer(data = hydrat_mod_dat_scaled,
                          osml_scaled ~ 
                           abshum_scaled*temp_scaled +
                            date + 
                            (1|individual_ID)) 
summary(hydrat_mod3)
```

I'm still getting the warning messages! aghhhhh

**@Dr. Bodwin please help with this ! D: **



### Best Models

```{r}
# save step 6 summary object
#osml_best_mod <- summary()
# extract stats table from summary object
#osml_best_mod_vals <- data.frame(osml_best_mod$coefficients)
# export 
#write.csv(osml_best_mod_vals, "osml_best_mod_vals.csv")
```


### Check LM Assumptions (Hydration Model )

First, get residuals:

```{r}
#res_hydrat_mod <- hydrat_mod_dat %>% # scaled or not? 
 # mutate(y_hat = predict(),
  #       e = residuals())
```


Linearity and Equal Variance

Is the function **linear**? Is there **equal** variance of the residuals? The residuals should be homoskedactic relative to y_hat (or x). We don't care if there is a relationship between the residuals ~ dependent variable (actual y).

Plotting residuals shows us whether the data meets linearity and equal variance assumptions:

```{r}
#ggplot(data = res_hydrat_mod, aes(x = y_hat, y = e)) +
 # geom_point() + 
  #theme_classic() + 
  #xlab("predicted y (y-hat)") +
  #ylab("residuals (e)") +
  #ggtitle("Hydration Model 5") +
  #geom_hline(yintercept = 0)
```


Brown-Forsythe test to statistically check equal variance:

H0: normally distributed (non-sig test is GOOD)
HA: NOT normally distributed (reject nul == assumption not satisfied)

```{r}
# need to create the right data & format first
#bf_data <- res_hydrat_mod %>%
#  dplyr::filter(complete.cases(temp_C_interpol)) %>%
#  dplyr::mutate(middle = median(temp_C_interpol),
#                side = temp_C_interpol > middle)
#bf_data$side <- as.factor(bf_data$side)

# now run test
#bf.test(formula = e ~ side, # y~x
 #       data = bf_data, # dataframe
  #      alpha = 0.05, # default 0.05
   #     na.rm = TRUE, # remove missing data before running?
    #    verbose = TRUE # print output to console?
     #   )
```

Equal variance is satisfied.


Now check normality. Is the distribution of residuals **normal**?

use Shapiro-Wilk normality test:
H0: data is NOT significantly different from normal distribution
HA: data IS significantly different from normal distribution

```{r}
#simple.eda(res_hydrat_mod$e)
#shapiro.test(res_hydrat_mod$e)
```


### Transformations 



### Conclusion



## CEWL

Based on the simple linear models and figures above, CEWL should be predicted by:
- body region
- cloacal temperature at measurement
- capture temperature, absolute humidity, wind speed, and solar radiation
- mass & SVL
- hold time (time between capture vs measurement)


Prep dataframe for models:

```{r}
CEWL_mod_dat <- all_data_long %>% 
  # select variables of interest
  dplyr::select(date,
                hold_time,
                individual_ID,
                mass_g,
                SVL_mm, 
                TEWL_g_m2h,
                region,
                cloacal_temp_C,
                temp_C_interpol,
                abs_humidity_g_m3_interpol,
                Wind_mph_interpol,
                Solar_rad_Wm2_interpol
                ) %>%
  dplyr::filter(complete.cases(.))
```


### Multicollinearity

Check for multicollinearity among independent variables:

```{r}
CEWL_mod_dat %>% 
  # get rid of dependent variable
  dplyr::select(-TEWL_g_m2h, -individual_ID) %>%
  # multicollinearity plot
  pairs(.)

# also make another plot with r-sq values
# non-numeric variables don't work for this
CEWL_mod_dat %>% 
  # select variables of interest
  dplyr::select(-TEWL_g_m2h, -date, -region, -individual_ID) %>% 
  # multicollinearity plot
  chart.Correlation(., histogram = F, pch = 19)
```

Mass and SVL are very collinear variables that should not be used in combination. Cloacal temp and hold time are pretty collinear, which makes sense; as we held them, they got cooler because they were inside and not basking. Individual ID and absolute humidity are also pretty collinear, but it's less intuitive. Temperature and solar radiation, as well as humidity and solar radiation, are each collinear pairs. We will use model selection to figure out which variable from each collinear pair is better to include in the model.



### Models & Selection

```{r}
# model 1
CEWL_mod1 <- lme4::lmer(data = CEWL_mod_dat,
                       TEWL_g_m2h ~ hold_time + 
                       region * mass_g + SVL_mm +
                       cloacal_temp_C +
                       temp_C_interpol +
                       abs_humidity_g_m3_interpol +
                       Wind_mph_interpol +
                       Solar_rad_Wm2_interpol + 
                       (1|individual_ID)) 
summary(CEWL_mod1)
```


Check which variables to drop:

```{r}
drop1(CEWL_mod1)
```

Based on AIC, dropping SVL, hold time, temp, humidity, and/or wind speed would result in a better model. 

Start with SVL and hold time:

```{r}
# model 2
CEWL_mod2 <- lme4::lmer(data = CEWL_mod_dat,
                       TEWL_g_m2h ~  
                       region * mass_g +
                       cloacal_temp_C +
                       temp_C_interpol +
                       abs_humidity_g_m3_interpol +
                       Wind_mph_interpol +
                       Solar_rad_Wm2_interpol + 
                       (1|individual_ID)) 
summary(CEWL_mod2)

# compare
anova(CEWL_mod2, CEWL_mod1)
```

AIC improved very slightly, but model 2 is not significantly better than model 1.

Check drop terms again:

```{r}
drop1(CEWL_mod2)
```


Temperature, humidity, and wind should still be deleted.

```{r}
# model 3
CEWL_mod3 <- lme4::lmer(data = CEWL_mod_dat,
                       TEWL_g_m2h ~
                       region * mass_g +
                       cloacal_temp_C +
                       Solar_rad_Wm2_interpol + 
                       (1|individual_ID)) 
summary(CEWL_mod3)

# compare
anova(CEWL_mod3, CEWL_mod1)
anova(CEWL_mod3, CEWL_mod2)
```

Once again, the AIC is slightly lower, but the model is not significantly better when compared to version 1 or 2. 

Technically, cloacal temperature should be a random effect. We can also check whether temp and humidity would be important as an interaction.

```{r}
# model 4
CEWL_mod4 <- lme4::lmer(data = CEWL_mod_dat,
                       TEWL_g_m2h ~
                       region * mass_g +
                       temp_C_interpol:abs_humidity_g_m3_interpol +
                       Solar_rad_Wm2_interpol + 
                       (1|individual_ID) + (1|cloacal_temp_C)) 
summary(CEWL_mod4)

# compare
anova(CEWL_mod3, CEWL_mod4)
```

Model 4 is significantly better than model 3. But, which change was more important? Check if we kept cloacal temperature as a fixed effect and added the temp:humidity interaction effect:

```{r}
# model 5
CEWL_mod5 <- lme4::lmer(data = CEWL_mod_dat,
                       TEWL_g_m2h ~
                       region * mass_g +
                       cloacal_temp_C +
                       temp_C_interpol:abs_humidity_g_m3_interpol +
                       Solar_rad_Wm2_interpol + 
                       (1|individual_ID)) 
summary(CEWL_mod5)

# compare
anova(CEWL_mod5, CEWL_mod4)
```

Model 4 is significantly better.

Compare model 4 to a new iteration without temp:humidity:

```{r}
# model 6
CEWL_mod6 <- lme4::lmer(data = CEWL_mod_dat,
                       TEWL_g_m2h ~
                       region * mass_g +
                       Solar_rad_Wm2_interpol + 
                       (1|individual_ID) + (1|cloacal_temp_C)) 
summary(CEWL_mod6)

# compare
anova(CEWL_mod6, CEWL_mod4)
```

Model 6 is not significantly better, and I think temp:humidity is an important variable to consider, so model 4 will be kept as our best model. 


### Best Model

The best model is CEWL predicted by:
- region * mass_g 
- temp_C_interpol:abs_humidity_g_m3_interpol 
- Solar_rad_Wm2_interpol
- individual_ID and cloacal_temp_C as random effects

```{r}
# save model 4 summary object
CEWL_best_mod <- summary(CEWL_mod4)
# extract stats table from summary object
CEWL_best_mod_vals <- data.frame(CEWL_best_mod$coefficients)
# export 
write.csv(CEWL_best_mod_vals, "CEWL_best_mod_vals.csv")
```



### Check LM Assumptions (CEWL Model 4)

First, get residuals:

```{r}
CEWL_mod_res <- CEWL_mod_dat %>%
  mutate(y_hat = predict(CEWL_mod4),
         e = residuals(CEWL_mod4))
```


Linearity and Equal Variance

Is the function **linear**? Is there **equal** variance of the residuals? The residuals should be homoskedactic relative to y_hat (or x). We don't care if there is a relationship between the residuals ~ dependent variable (actual y).

Plotting residuals shows us whether the data meets linearity and equal variance assumptions:

```{r}
ggplot(data = CEWL_mod_res, aes(x = y_hat, y = e)) +
  geom_point() + 
  theme_classic() + 
  xlab("predicted y (y-hat)") +
  ylab("residuals (e)") +
  ggtitle("CEWL Model 13 Residuals") +
  geom_hline(yintercept = 0)
```

It's definitely making a fan shape. :(

Brown-Forsythe test to statistically check equal variance:

H0: normally distributed (non-sig test is GOOD)
HA: NOT normally distributed (reject nul == assumption not satisfied)

```{r}
# need to create the right data & format first
bf_data <- CEWL_mod_res %>%
  dplyr::mutate(middle_mass = median(mass_g), # mass
                side_mass = as.factor(mass_g > middle_mass),
                # solar radiation
                middle_sorad = median(Solar_rad_Wm2_interpol),
                side_sorad = as.factor(Solar_rad_Wm2_interpol > middle_sorad), 
                # temperature
                middle_temp = median(temp_C_interpol),
                side_temp = as.factor(temp_C_interpol > middle_temp),
                # absolute humidity
                middle_absh = median(Solar_rad_Wm2_interpol),
                side_absh = as.factor(Solar_rad_Wm2_interpol > middle_absh)
                )

# now run test
bf.test(formula = e ~ side_absh, # y~x
        data = bf_data, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
bf.test(formula = e ~ side_temp, # y~x
        data = bf_data, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
bf.test(formula = e ~ side_sorad, # y~x
        data = bf_data, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
bf.test(formula = e ~ side_mass, # y~x
        data = bf_data, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
```

Equal variance is satisfied for all 4 continuous predictor variables.


Now check normality. Is the distribution of residuals **normal**?

use Shapiro-Wilk normality test:
H0: data is NOT significantly different from normal distribution
HA: data IS significantly different from normal distribution

```{r}
simple.eda(CEWL_mod_res$e)
shapiro.test(CEWL_mod_res$e)
```

not normal!


### Test Transformations 

Can I improve satisfaction of LM assumptions by transforming the dependent variable?

```{r}
CEWL_transf <- all_data_long %>%
  mutate(TEWL_sqrt = sqrt(TEWL_g_m2h),
         TEWL_log = log(TEWL_g_m2h))

# sqrt(TEWL)
CEWL_transf %>%
  ggplot(., aes(x = TEWL_sqrt)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("sqrt(CEWL)") + 
  ylab("Count") + 
  facet_wrap(~region)

# log(temperature)
CEWL_transf %>%
  ggplot(., aes(x = TEWL_log)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("LOG(CEWL)") + 
  ylab("Count") + 
  facet_wrap(~region)
```


Log transforming seems to be pretty effective across body regions.



### Transform & Re-Model

I will log-transform CEWL and see whether it makes the models satisfy LMM assumptions better.

Run CEWL model 4 with log-transformed CEWL:

```{r}
# log-transformed model 4
CEWL_mod4_t <- lme4::lmer(data = CEWL_mod_dat,
                       log(TEWL_g_m2h) ~
                       region * mass_g +
                       temp_C_interpol:abs_humidity_g_m3_interpol +
                       Solar_rad_Wm2_interpol + 
                       (1|individual_ID) + (1|cloacal_temp_C)) 
summary(CEWL_mod4_t)

# compare
anova(CEWL_mod4_t, CEWL_mod4)
```

The transformed model is WAYYYY better! :D


### Re-Check Assumptions (transformed model 4)

First, get residuals:

```{r}
CEWL_t_mod_res <- CEWL_mod_dat %>%
  mutate(y_hat = predict(CEWL_mod4_t),
         e = residuals(CEWL_mod4_t))
```


Linearity and Equal Variance

Is the function **linear**? Is there **equal** variance of the residuals? The residuals should be homoskedactic relative to y_hat (or x). We don't care if there is a relationship between the residuals ~ dependent variable (actual y).

Plotting residuals shows us whether the data meets linearity and equal variance assumptions:

```{r}
ggplot(data = CEWL_t_mod_res, aes(x = y_hat, y = e)) +
  geom_point() + 
  theme_classic() + 
  xlab("predicted y (y-hat)") +
  ylab("residuals (e)") +
  ggtitle("CEWL Model 13 Residuals") +
  geom_hline(yintercept = 0)
```

It looks much much better. :)

Brown-Forsythe test to statistically check equal variance:

H0: normally distributed (non-sig test is GOOD)
HA: NOT normally distributed (reject nul == assumption not satisfied)

```{r}
# need to create the right data & format first
bf_data_t <- CEWL_t_mod_res %>%
  dplyr::mutate(middle_mass = median(mass_g), # mass
                side_mass = as.factor(mass_g > middle_mass),
                # solar radiation
                middle_sorad = median(Solar_rad_Wm2_interpol),
                side_sorad = as.factor(Solar_rad_Wm2_interpol > middle_sorad), 
                # temperature
                middle_temp = median(temp_C_interpol),
                side_temp = as.factor(temp_C_interpol > middle_temp),
                # absolute humidity
                middle_absh = median(Solar_rad_Wm2_interpol),
                side_absh = as.factor(Solar_rad_Wm2_interpol > middle_absh)
                )

# now run test
bf.test(formula = e ~ side_absh, # y~x
        data = bf_data_t, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
bf.test(formula = e ~ side_temp, # y~x
        data = bf_data_t, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
bf.test(formula = e ~ side_sorad, # y~x
        data = bf_data_t, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
bf.test(formula = e ~ side_mass, # y~x
        data = bf_data_t, # dataframe
        alpha = 0.05, # default 0.05
        na.rm = TRUE, # remove missing data before running?
        verbose = TRUE # print output to console?
        )
```

Equal variance is still satisfied for all 4 continuous predictor variables.


Now check normality. Is the distribution of residuals **normal**?

use Shapiro-Wilk normality test:
H0: data is NOT significantly different from normal distribution
HA: data IS significantly different from normal distribution

```{r}
simple.eda(CEWL_t_mod_res$e)
shapiro.test(CEWL_t_mod_res$e)
```

Still not statistically normal... but the distribution looks a lot better.


### Conclusion

The best CEWL model should use log-transformed CEWL because this greatly improves the model based on AIC and it allows the model to satisfy the linearity assumption of LMM. 

```{r}
# save model 4 summary object
CEWL_best_t_mod <- summary(CEWL_mod4_t)
# extract stats table from summary object
CEWL_best_t_mod_vals <- data.frame(CEWL_best_t_mod$coefficients)
# export 
write.csv(CEWL_best_t_mod_vals, "CEWL_best_mod_vals.csv")
```





# What to Present in the Paper

- figures (exported)
- hct SLR
- best osml mod (transformed?)
- best CEWL mod (CEWL transformed)

