---
title: "Sceloporus Humidity Acclimation 2021 - Experimental Data Analysis"
author: "Savannah Weaver"
output: 
  pdf_document:
    toc: TRUE
---

# Packages

```{r setup, include=FALSE}
`%nin%` = Negate(`%in%`)
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse")
if (!require("UsingR")) install.packages("UsingR")
library("UsingR")
if (!require("lme4")) install.packages("lme4")
library("lme4") # for LMMs
if (!require("lmerTest")) install.packages("lmerTest")
library("lmerTest") # for p-values
if (!require("broom")) install.packages("broom")
library("broom") # lmer model export
if (!require("broom.mixed")) install.packages("broom.mixed")
library("broom.mixed") # lmer model export
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # multi-ggplot figures
if (!require("AICcmodavg")) install.packages("AICcmodavg")
library("AICcmodavg") # VIF calculations
if (!require("car")) install.packages("car")
library("car") # VIF calculations
```




# Data

This data was collected in the Spring of 2021 in conjunction with a study carried out in Cal Poly's Herpetology class. Some lizards measured for that primary study were kept to observe physiological changes in response to different climate treatments. See (doi) for full details


## Morphometrics & Hydration

### Treatment Groups

variables:
- individual lizard ID
- temp_tmt_C = temperature treatment
- humidity_tmt_percent = humidity treatment (high/low, not actually %)
- trial_number = which set of lizards that individual was from
- conclusion = how that individual's experiment ended (died, canceled, or complete)

```{r tmt data}
tmts <- read.csv("./data/exp_tmt_assignment.csv")
```


### Capture Data

variables:
- date = date of capture & baseline measurements
- individual lizard ID
- mass_g = mass in grams
- hematocrit_percent = % of blood sample that's red blood cells
- osmolality_mmol_kg = concentration of solutes in blood plasma
- type = when the measurements were taken along the course of the experiment (all on capture day)

```{r capture data}
capture_hydration <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  mutate(# correctly format date-only variable
         date = as.Date(date, format = "%Y-%m-%d")
         ) %>%
  # select only relevant variables
  dplyr::select(date, individual_ID, 
                mass_g, hematocrit_percent, osmolality_mmol_kg
                ) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID) %>%
  mutate(type = as.factor("capture"))
summary(capture_hydration)
```


extract SVL data separately from capture data:

```{r SVL data}
SVL <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::select(individual_ID, SVL_mm) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID)
summary(SVL)
```


extract capture CEWL cloacal temperature separately:

```{r cap CT data}
cap_CT <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::select(individual_ID, cloacal_temp_C) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID)
summary(cap_CT)
```



### Experiment Data

variables:
- date = date of measurements
- individual lizard ID
- mass_g = mass in grams
- hematocrit_percent = % of blood sample that's red blood cells
- osmolality_mmol_kg = concentration of solutes in blood plasma (mean of 1-3 replicates)
- type = when the measurements were taken along the course of the experiment (either during experimental treatment or after rehab)

```{r exp data}
exp_dat <- read.csv("./data/experimental_data.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
                # format date
  dplyr::mutate(date = as.Date(date, format = "%m/%d/%y"),
                type = as.factor(type)
                ) %>%
  # select only variables to be analyzed
  dplyr::select(date, individual_ID, mass_g, 
                hematocrit_percent, type, 
                osmolality_mmol_kg = osmolality_mmol_kg_replicate_mean)
summary(exp_dat)
```

### Join Dataframes

Now, attach all the dataframes, only use individuals whose treatment was completed, and add a "day" variable for what day of treatment each lizard/observation was on. I also calculate SMI using the equation created in capture_analysis.


```{r join morpho blood data}

all_dat <- exp_dat %>%
  # join data
  rbind(capture_hydration) %>%
  # add tmt group info
  left_join(tmts, by = "individual_ID") %>%
  dplyr::select(-notes) %>%
  # add SVL value for each obs of each indiv.
  # for computing BCI and scaled mass indices
  left_join(SVL, by = "individual_ID") %>%
  # only use completed experiment runs
  dplyr::filter(conclusion == "complete") %>%
  group_by(individual_ID) %>%
  # reformat a lot of variables
  mutate(capture_date = min(date),
         day = as.numeric(date - capture_date),
         humidity_tmt_percent = as.factor(humidity_tmt_percent),
         individual_ID = as.factor(individual_ID),
         temp_tmt_C = as.factor(temp_tmt_C),
         trial_number = as.factor(trial_number),
         conclusion = as.factor(conclusion),
         SMI = mass_g * ((65.02158/SVL_mm) ^ (3.09059/sqrt(0.8944)))
         ) %>%
  # in the first trial we took measurements every 2 days
  # exclude those obs to make consistent across trials
  dplyr::filter(day %nin% c(2,6))

summary(all_dat)
unique(all_dat$individual_ID)
```

re-order some factors:

```{r reorder all_dat humidity factor}
all_dat$humidity_tmt_percent <- factor(all_dat$humidity_tmt_percent,
                                       levels = c("humid", "dry"),
                                       labels = c("Humid", "Dry"))
```


```{r}
all_dat$day <- factor(all_dat$day,
                      levels = c("0", "4", "8", "9", "10", "11"),
                      labels = c("Before Acclimation", 
                                 "Mid Acclimation", 
                                 "After Acclimation", 
                                 "After Acclimation",
                                 "After Rehydration",
                                 "After Rehydration"))
summary(all_dat)
```


make a sub-dataframe without rehab data to prevent any mix-ups:

```{r extra df}
all_dat_no_rehab <- all_dat %>%
  dplyr::filter(type != "rehab")
```


### Checks

Dates:

```{r check dates}
# check that capture dates are valid
unique(all_dat$capture_date)
```

Check that each lizard has an accurate number of measurements.

```{r check n measures per lizard}
all_dat %>%
  group_by(individual_ID, type) %>%
  summarise(n = n()) %>%
  arrange(type)
```


That all looks good, every lizard has 1 capture measurement, 2 experimental measurements, and 1 rehab measurement.


## CEWL

### Capture CEWL

variables:
- date = date of capture & baseline measurements
- individual lizard ID
- region = which body area the measurement was taken from
- TEWL_g_m2h = evaporative water loss
- cloacal_temp_C = taken at measurement; influences CEWL

```{r cap CEWL data}
cap_CEWL <- read.csv("./exported_data/capture_CEWL.csv") %>%
  dplyr::select(date, individual_ID, region, TEWL_g_m2h) %>%
  mutate(#individual_ID = as.factor(individual_ID), # do later
         date = as.Date(date, format = "%Y-%m-%d"),
         region = as.factor(region),
         day = as.factor("before"),
         n_day = 0
         ) %>%
  dplyr::filter(individual_ID %in% all_dat$individual_ID) %>%
  left_join(cap_CT, by = 'individual_ID')
summary(cap_CEWL)
```


### Post-Experiment CEWL

Load in each of the post-rehab datafiles:

```{r exp CEWL data}
# trial 1
CEWL_t1 <- read.csv("./data/post_exp_CEWL/4-28-21-CEWL.csv", # filename
                          na.strings=c("","NA")) %>% # fix empty cells
  # rename and select the pertinent variables/cols
  # I have to do this for each one 
  # so they all have the same number of columns for joining
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.. # rename
                )

# trial 2
CEWL_t2 <- read.csv("./data/post_exp_CEWL/5-4-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )

# trial 3
CEWL_t3 <- read.csv("./data/post_exp_CEWL/5-11-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )

# trial 4
CEWL_t4 <- read.csv("./data/post_exp_CEWL/5-18-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )
```


Load in cloacal temperatures:

```{r exp CT data}
exp_CT <- read.csv("./data/post_exp_CEWL_cloacal_temps.csv") %>%
  mutate(date = as.Date(date, format = "%Y/%m/%d")) %>%
  dplyr::select(-time)
summary(exp_CT)
```


### Join Dataframes 

Merge all post-experiment CEWL, add cloacal temperature, add capture CEWL:

```{r join CEWL data}
# merge all CEWL datafiles & reformat
CEWL <- CEWL_t1 %>% # trial 1
  rbind(., CEWL_t2, # trial 2
        CEWL_t3, # trial 3
        CEWL_t4 # trial 4
        ) %>%
  # remove any unsuccessful measurements
  dplyr::filter(Status == "Normal") %>%
  # extract individual_ID and region separately from the "ID" variable
  separate(ID, c("individual_ID", "region")) %>%
  # reformat data
  dplyr::mutate(# reformat date
                date = as.Date(date, format = "%m/%d/%y"),
                # format individual ID
                individual_ID = as.integer(individual_ID),
                # set body region as a factor variable after getting only the consistent characters due to typos
                region = as.factor(substring(region, 1, 4)),
                # add when measurement taken
                day = as.factor("after"),
                n_day = 1 # technically day 8/9, just to help with figures
                ) %>%
  # remove cols not relevant to stats
  dplyr::select(-Status) %>%
  # remove any rows with missing values
  # none actually needed to be removed
  dplyr::filter(complete.cases(.)) %>%
  # add cloacal temperatures
  left_join(exp_CT, by = c("date", "individual_ID")) %>%
  # now matching dataframes, add capture CEWL data
  rbind(cap_CEWL) %>%
  # add tmt assignments
  left_join(tmts, by = "individual_ID") %>%
  mutate(humidity_tmt_percent = as.factor(humidity_tmt_percent),
         individual_ID = as.factor(individual_ID),
         conclusion = as.factor(conclusion),
         trial_number = as.factor(trial_number)
         ) %>%
  # lizards 49 & 80 are missing pre-exp CEWL, so remove them
  dplyr::filter((individual_ID %nin% c('49', '80')))
# every lizard should have 10 measurements
summary(CEWL)
```

Check that data looks correct:

```{r}
CEWL %>%
  group_by(individual_ID, day) %>%
  summarise(n = n()) %>%
  arrange(individual_ID, n)
```

Everything looks great! (after removing the observations for the two lizards with missing pre-experiment CEWL measurements.)

Before/after aren't perfectly even because sometimes we were unable to get the AquaFlux to equilibrate and take a measurement.

Finally, make a small edit so the regions are spelled out completely. This requires reordering factor levels:

```{r reorder CEWL factors}
CEWL$region <- factor(CEWL$region, 
                      levels = c("dors", "vent", "head", "dewl", "mite"),
                      labels = c("Dorsum", "Ventrum", "Head", 
                                 "Dewlap", "Mite Patch")
                  )
CEWL$humidity_tmt_percent <- factor(CEWL$humidity_tmt_percent,
                                       levels = c("humid", "dry"),
                                       labels = c("Humid", "Dry"))
CEWL$day <- factor(CEWL$day,
                                       levels = c("before", "after"),
                                       labels = c("Before", "After"))
summary(CEWL)
```



## Export Data Frames for Power Analyses

```{r write data}
#write.csv(all_dat, "exported_data/exp_effects_hydration.csv")
#write.csv(CEWL, "exported_data/exp_effects_CEWL.csv")
```



# Figures 

## Means to Overlay

```{r}
all_dat_mean_SMI <- all_dat %>%
  group_by(humidity_tmt_percent, day) %>%
  summarise(SMI_mean = mean(SMI))
all_dat_mean_SMI
all_dat_mean_hct <- all_dat %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  group_by(humidity_tmt_percent, day) %>%
  summarise(hct_mean = mean(hematocrit_percent))
all_dat_mean_hct
all_dat_mean_osml <- all_dat %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>%
  group_by(humidity_tmt_percent, day) %>%
  summarise(osml_mean = mean(osmolality_mmol_kg))
all_dat_mean_osml
all_dat_mean_CEWL <- read.csv("./data/CEWL_predicted_exp_means.csv") %>%
  mutate(region = as.factor(region),
         humidity_tmt_percent = as.factor(humidity_tmt_percent))
all_dat_mean_CEWL
```


## SMI ~ Time

```{r SMI fig}
ggplot() + 
  geom_point(data = all_dat,
             aes(x = day,
                 y = SMI, 
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  geom_line(data = all_dat,
            aes(x = day,
                y = SMI,
                group = individual_ID,
                color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                ),
            alpha = 0.4) +
  geom_line(data = all_dat_mean_SMI,
            aes(x = day,
                y = SMI_mean,
                group = humidity_tmt_percent,
                color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                ),
            size = 1.5,
            alpha = 1) +
  geom_point(data = all_dat_mean_SMI,
             aes(x = day,
                 y = SMI_mean,
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent),
             size = 5, 
             alpha = 1) +
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = "") +
  scale_shape_manual(values = c(17, 19), name = "") +
  #scale_linetype_manual(values = c("solid", "dotdash"),
   #                  name = "") +
  xlab("") + 
  ylab("Body Condition\n(g)") + 
  geom_vline(xintercept = 3.5,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 26),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 18),
        axis.text.x = element_blank(),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 28),
        legend.text.align = 0,
        legend.position = "none",
        plot.margin = unit(c(0.1, #top
                             0.1, #right
                             0.1, #bottom
                             0.46 #left
                             ), "cm")
) -> tmt_effects_SMI
tmt_effects_SMI

# export figure
#ggsave(filename = "tmt_effects_SMI.jpeg",
 #      plot = tmt_effects_SMI,
  #     path = "./final_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 5, height = 4)
```


## Hct ~ Time

```{r hct fig}
ggplot() + 
  geom_point(data = all_dat, 
             aes(x = day,
                 y = hematocrit_percent, 
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  geom_line(data = all_dat,
            aes(x = day,
                y = hematocrit_percent,
                group = individual_ID,
                color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                ),
            alpha = 0.4) +
  geom_line(data = all_dat_mean_hct,
            aes(x = day,
                y = hct_mean,
                group = humidity_tmt_percent,
                color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                ),
            size = 1.5,
            alpha = 1) +
  geom_point(data = all_dat_mean_hct,
             aes(x = day,
                 y = hct_mean,
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent),
             size = 5, 
             alpha = 1) +
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = "") +
  scale_shape_manual(values = c(17, 19), name = "") +
  #scale_linetype_manual(values = c("solid", "dotdash"),
   #                  name = "") +
  scale_x_discrete(labels = c("Before\nAcclimation", 
                              "Mid\nAcclimation",
                              "After\nAcclimation",
                              "After\nRehydration")) +
  xlab("") + 
  ylab("Hematocrit\n(%)") + 
  geom_vline(xintercept = 3.5,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 26),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 18),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 28),
        legend.text.align = 0,
        legend.position = "none",
        plot.margin = unit(c(0.1, #top
                             0.1, #right
                             0.1, #bottom
                             0.46 #left
                             ), "cm")
        ) -> tmt_effects_hct
tmt_effects_hct

# export figure
#ggsave(filename = "tmt_effects_hct.jpeg",
 #      plot = tmt_effects_hct,
  #     path = "./final_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 5, height = 4)
```



## Osml ~ Time

```{r osmolality fig}
ggplot() + 
  geom_point(data = all_dat,
             aes(x = day,
                 y = osmolality_mmol_kg, 
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  geom_line(data = all_dat,
            aes(x = day,
                y = osmolality_mmol_kg,
                group = individual_ID,
                color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                ),
            alpha = 0.4) +
  geom_line(data = all_dat_mean_osml,
            aes(x = day,
                y = osml_mean,
                group = humidity_tmt_percent,
                color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                ),
            size = 1.5,
            alpha = 1) +
  geom_point(data = all_dat_mean_osml,
             aes(x = day,
                 y = osml_mean,
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent),
             size = 5, 
             alpha = 1) +
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = "") +
  scale_shape_manual(values = c(17, 19), name = "") +
  #scale_linetype_manual(values = c("solid", "dotdash"),
   #                  name = "") +
  xlab("") + 
  ylab("Plasma Osmolality\n(mmol/kg)") + 
  geom_vline(xintercept = 3.5,
             linetype = "dashed",
             color = "darkgrey") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 26),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 18),
        axis.text.x = element_blank(),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 28),
        legend.text.align = 0,
        legend.position = "none",
        plot.margin = unit(c(0.1, #top
                             0.1, #right
                             0.1, #bottom
                             0.1 #left
                             ), "cm")
        ) -> tmt_effects_osml
tmt_effects_osml

# export figure
#ggsave(filename = "tmt_effects_osml.jpeg",
 #      plot = tmt_effects_osml,
  #     path = "./final_figures",
   #    device = "jpeg",
    #   dpi = 1200,
     #  width = 5, height = 4)
```


## Multi-Figure

```{r}
ggarrange(tmt_effects_SMI, tmt_effects_osml, tmt_effects_hct,
          ncol = 1, nrow = 3,
          labels = c("A", "B", "C"),
          font.label = list(size = 24, face = "bold", color ="black"),
          hjust = -2, vjust = 1.1,
          common.legend = TRUE,
          legend = "bottom"
          ) -> tmt_multi_fig
tmt_multi_fig
# export figure
ggsave(filename = "tmt_multi_fig2.jpeg",
       plot = tmt_multi_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 8, height = 12)
```




## CEWL

```{r CEWL fig}
ggplot(data = CEWL) + 
  geom_point(aes(x = n_day,
                 y = TEWL_g_m2h, 
                 group = individual_ID,
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  geom_line(aes(x = n_day,
                y = TEWL_g_m2h,
                group = individual_ID,
                color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                ),
            alpha = 0.4) +
  stat_smooth(aes(x = n_day, 
                  y = TEWL_g_m2h, 
                  color = humidity_tmt_percent,
                #linetype = humidity_tmt_percent
                  ),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.5, 
              alpha = 1) + 
  geom_point(data = all_dat_mean_CEWL,
             aes(x = n_day,
                 y = CEWL_mean, 
                 group = humidity_tmt_percent,
                 color = humidity_tmt_percent,
                 shape = humidity_tmt_percent
                 ), 
             size = 5, 
             alpha = 1) +
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = ""
                     #name = "Humidity\nTreatment"
                     ) +
  scale_shape_manual(values = c(17, 19), name = "") +
  #scale_linetype_manual(values = c("solid", "dotdash"),
   #                     name = "") +
  facet_wrap(~region, ncol = 2) +
  scale_x_continuous(breaks = c(0, 1),
                     labels = c("0" = "Before\nAcclimation", 
                                "1" = "After\nAcclimation")
                     ) +
  xlab("") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 24),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 18),
        axis.text.x = element_text(#angle = 90,
                                   hjust = c(0, 1),
                                   #vjust = 0.2
                                   ),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 26),
        legend.text.align = 0,
        legend.position = c(0.75,0.12),
        #legend.justification = c(1, 1)
) -> CEWL_tmt_fig
CEWL_tmt_fig

# export figure
ggsave(filename = "tmt_effects_CEWL.jpeg",
       plot = CEWL_tmt_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 8, height = 12)
```







# Models


## SMI



Check whether means started out different:

```{r SMI check original difference}
SMI_diff_lm <- all_dat_no_rehab %>%
  dplyr::filter(day == "Before Acclimation") %>%
  lm(data = ., SMI ~ humidity_tmt_percent)
summary(SMI_diff_lm)
```

NOT significantly different, which is good.

Check whether means ended differently:

```{r SMI check ending difference}
SMI_diff_lm_end <- all_dat_no_rehab %>%
  dplyr::filter(day == "After Acclimation") %>%
  lm(data = ., SMI ~ humidity_tmt_percent)
summary(SMI_diff_lm_end)
```


### Build Model

```{r SMI models}
SMI_mod1 <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ day*humidity_tmt_percent +
               (1|trial_number/individual_ID))
drop1(SMI_mod1)

# drop interaction term
SMI_mod2 <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ day + humidity_tmt_percent +
               (1|trial_number/individual_ID))
drop1(SMI_mod2)

# drop humidity effect
SMI_mod3 <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ day +
               (1|trial_number/individual_ID))
drop1(SMI_mod3)

# null model
SMI_mod_null <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ 1 +
               (1|trial_number/individual_ID))
```


### Selection

```{r SMI mod compare}
SMI_models <- list(SMI_mod1, SMI_mod2, SMI_mod3, SMI_mod_null)

#specify model names
SMI_mod_names <- c('(model 1) ~ day * humidity', 
                   '(model 2) ~ day + humidity', 
                   '(model 3) ~ day',
                   'null model')
#calculate AIC of each model
SMI_AICc <- data.frame(aictab(cand.set = SMI_models, 
                                 modnames = SMI_mod_names))
SMI_AICc
```


### Check Conditions of Top Models


```{r}
plot(SMI_mod2)
plot(SMI_mod3)
car::vif(SMI_mod2)
```

L & E both look good.


Is the distribution of residuals **normal**?
use Shapiro-Wilk normality test:
H0: data is NOT significantly different from normal distribution
HA: data IS significantly different from normal distribution

```{r}
simple.eda(residuals(SMI_mod2))
shapiro.test(residuals(SMI_mod2))
simple.eda(residuals(SMI_mod3))
shapiro.test(residuals(SMI_mod3))
```

Normality is fine.

Re-run top 2 models with lmerTest for p-values:

```{r SMI top mod ps}
SMI_mod2p <- lmerTest::lmer(data = all_dat_no_rehab,
               SMI ~ day + humidity_tmt_percent +
               (1|trial_number/individual_ID))
summary(SMI_mod2p)
SMI_mod3p <- lmerTest::lmer(data = all_dat_no_rehab,
               SMI ~ day +
               (1|trial_number/individual_ID))
summary(SMI_mod2p)
```



### Export

Save information for the top models, as well as the table showing the rankings of the different models.

```{r save SMI}
write.csv(SMI_AICc, "./best_models/exp_effects_SMI_mod_rankings.csv")
write.csv(broom.mixed::tidy(SMI_mod2p), 
          "./best_models/exp_effects_SMI_best_mod1.csv")
write.csv(broom.mixed::tidy(SMI_mod3p), 
          "./best_models/exp_effects_SMI_best_mod2.csv")
```




## Hematocrit

### Build Model

```{r hct models}
hct_mod1 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ day * humidity_tmt_percent +
               (1|trial_number/individual_ID))
summary(hct_mod1)
drop1(hct_mod1)

# drop day*humidity interaction
hct_mod2 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ day + humidity_tmt_percent +
               (1|trial_number/individual_ID))
summary(hct_mod2)
drop1(hct_mod2)

# drop humidity tmt
hct_mod3 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ day +
               (1|trial_number/individual_ID))
summary(hct_mod3)
drop1(hct_mod3)

# null model
hct_mod_null <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ 1 +
               (1|trial_number/individual_ID))
summary(hct_mod_null)
```


### Selection

```{r hct mod compare}
hct_models <- list(hct_mod1, hct_mod2, hct_mod3, hct_mod_null)

#specify model names
hct_mod_names <- c('(model 1) ~ day * humidity', 
                   '(model 2) ~ day + humidity', 
                   '(model 3) ~ day',
                   'null model')
#calculate AIC of each model
hct_AICc <- data.frame(aictab(cand.set = hct_models, 
                                 modnames = hct_mod_names))
hct_AICc
```


### Check Conditions of Top Model


```{r}
plot(hct_mod1)
car::vif(hct_mod1)
```

looks okay

Is the distribution of residuals **normal**?
use Shapiro-Wilk normality test:
H0: data is NOT significantly different from normal distribution
HA: data IS significantly different from normal distribution

```{r}
simple.eda(residuals(hct_mod1))
shapiro.test(residuals(hct_mod1))
```

residuals are normally distributed!


Re-run top model with lmerTest for p-values:

```{r hct top mod ps}
hct_mod1p <- lmerTest::lmer(data = all_dat_no_rehab,
               hematocrit_percent ~ day * humidity_tmt_percent +
               (1|trial_number/individual_ID))
summary(hct_mod1p)
```



### Export 

```{r save hct}
write.csv(hct_AICc, "./best_models/exp_effects_hct_mod_rankings.csv")
write.csv(broom.mixed::tidy(hct_mod1p), 
          "./best_models/exp_effects_hct_best_mod1.csv")
```


## Osmolality


### Build Model

```{r osml models}
osml_mod1 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>%
  lme4::lmer(data = .,
               osmolality_mmol_kg ~ day * humidity_tmt_percent +
               (1|trial_number/individual_ID))
summary(osml_mod1)
drop1(osml_mod1)

# drop day*humidity interaction
osml_mod2 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>%
  lme4::lmer(data = .,
               osmolality_mmol_kg ~ day + humidity_tmt_percent +
               (1|trial_number/individual_ID))
summary(osml_mod2)
drop1(osml_mod2)

# drop humidity tmt
osml_mod3 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>%
  lme4::lmer(data = .,
               osmolality_mmol_kg ~ day +
               (1|trial_number/individual_ID))
summary(osml_mod3)
drop1(osml_mod3)

# null model
osml_mod_null <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>%
  lme4::lmer(data = .,
               osmolality_mmol_kg ~ 1 +
               (1|trial_number/individual_ID))
summary(osml_mod_null)
```


### Selection

```{r osml mod compare}
osml_models <- list(osml_mod1, osml_mod2, osml_mod3, osml_mod_null)

#specify model names
osml_mod_names <- c('(model 1) ~ day * humidity', 
                   '(model 2) ~ day + humidity', 
                   '(model 3) ~ day',
                   'null model')
#calculate AIC of each model
osml_AICc <- data.frame(aictab(cand.set = osml_models, 
                                 modnames = osml_mod_names))
osml_AICc
```


### Check Conditions of Top Model


```{r}
plot(osml_mod1)
car::vif(osml_mod1)
```

looks pretty good

Is the distribution of residuals **normal**?
use Shapiro-Wilk normality test:
H0: data is NOT significantly different from normal distribution
HA: data IS significantly different from normal distribution

```{r}
simple.eda(residuals(osml_mod1))
shapiro.test(residuals(osml_mod1))
```

residuals are normally distributed!


Re-run top model with lmerTest for p-values:

```{r osml top mod ps}
osml_mod1p <- lmerTest::lmer(data = all_dat_no_rehab,
               osmolality_mmol_kg ~ day * humidity_tmt_percent +
               (1|trial_number/individual_ID))
summary(osml_mod1p)
```



### Export 

```{r save osml}
write.csv(osml_AICc, "./best_models/exp_effects_osml_mod_rankings.csv")
write.csv(broom.mixed::tidy(osml_mod1p), 
          "./best_models/exp_effects_osml_best_mod1.csv")
```




## CEWL

### Build Model

```{r CEWL models}
# FULL model
CEWL_mod1 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ day * humidity_tmt_percent * region +
               cloacal_temp_C +
               (1|trial_number/individual_ID))
summary(CEWL_mod1)
drop1(CEWL_mod1)

# drop triple interaction
CEWL_mod2 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ day * (humidity_tmt_percent + region) +
                 humidity_tmt_percent * region +
               cloacal_temp_C +
               (1|trial_number/individual_ID))
summary(CEWL_mod2)
drop1(CEWL_mod2)

# drop humidity*region interaction
CEWL_mod3 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ day * (humidity_tmt_percent + region) +
               cloacal_temp_C +
               (1|trial_number/individual_ID))
summary(CEWL_mod3)
drop1(CEWL_mod3)

# drop day*region interaction
CEWL_mod4 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ day * humidity_tmt_percent + region +
               cloacal_temp_C +
               (1|trial_number/individual_ID))
summary(CEWL_mod4)
drop1(CEWL_mod4)

# drop cloacal temp
CEWL_mod5 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ day * humidity_tmt_percent + region +
               (1|trial_number/individual_ID))
summary(CEWL_mod5)
drop1(CEWL_mod5)

# drop day*humidity interaction
CEWL_mod6 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ day + humidity_tmt_percent + region +
               (1|trial_number/individual_ID))
summary(CEWL_mod6)
drop1(CEWL_mod6)

# drop humidity
CEWL_mod7 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ day + region +
               (1|trial_number/individual_ID))
summary(CEWL_mod7)
drop1(CEWL_mod7)

# drop day
CEWL_mod8 <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ region +
               (1|trial_number/individual_ID))
summary(CEWL_mod8)
drop1(CEWL_mod8)

# NULL model
CEWL_mod_null <- lme4::lmer(data = CEWL,
               TEWL_g_m2h ~ 1 +
               (1|trial_number/individual_ID))
```




### Selection

```{r CEWL mod compare}
CEWL_models <- list(CEWL_mod1, CEWL_mod2, CEWL_mod3, CEWL_mod4, 
                    CEWL_mod5, CEWL_mod6, CEWL_mod7, CEWL_mod8,
                    CEWL_mod_null)

#specify model names
CEWL_mod_names <- c('(model 1) ~ day * humidity * region + cloacal temp', 
                    '(model 2) ~ day * humidity + day * region + humidity * region + cloacal temp', 
                    '(model 3) ~ day * humidity + day * region + cloacal temp',
                    '(model 4) ~ day * humidity + region + cloacal temp',
                    '(model 5) ~ day * humidity + region',
                    '(model 6) ~ day + humidity + region',
                    '(model 7) ~ day + region',
                    '(model 8) ~ region',
                    'null model')
#calculate AIC of each model
CEWL_AICc <- data.frame(aictab(cand.set = CEWL_models, 
                                 modnames = CEWL_mod_names))
CEWL_AICc
```

The FULL model is the best model.


### Predict

Use the model to get the predicted means, which I will go back and use in the tmt_effects_CEWL figure.

```{r}
CEWL_predicts <- CEWL %>%
  mutate(pred = predict(CEWL_mod1)) %>%
  group_by(region, humidity_tmt_percent, n_day) %>%
  summarise(CEWL_mean = mean(pred))
CEWL_predicts
#write.csv(CEWL_predicts, "./data/CEWL_predicted_exp_means.csv")
```


### Check Conditions

Is the function **linear**? Is there **equal** variance of the residuals? The residuals should be homoskedactic relative to y_hat (or x). Plotting residuals shows us whether the data meets linearity and equal variance assumptions:

```{r}
plot(CEWL_mod1)
car::vif(CEWL_mod1)
```

clear fanning shape... L & E are not satisfied.


### Transform CEWL

Log-transformation worked well in the capture analysis, so try that here.

```{r CEWL mod transform}
CEWL_mod1t <- lme4::lmer(data = CEWL,
               log(TEWL_g_m2h) ~ day * humidity_tmt_percent * region +
               cloacal_temp_C +
               (1|trial_number/individual_ID))
plot(CEWL_mod1t)
simple.eda(residuals(CEWL_mod1t))
shapiro.test(residuals(CEWL_mod1t))
```

L, E, and N are all satisfied now. :)



Re-run top model with lmerTest for p-values:

```{r CEWL top mod ps}
CEWL_mod1tp <- lmerTest::lmer(data = CEWL,
               log(TEWL_g_m2h) ~ day * humidity_tmt_percent * region +
               cloacal_temp_C +
               (1|trial_number/individual_ID))
summary(CEWL_mod1tp)
```


### Export 


```{r save CEWL}
write.csv(CEWL_AICc, "./best_models/exp_effects_CEWL_mod_rankings.csv")
write.csv(broom.mixed::tidy(CEWL_mod1tp), 
          "./best_models/exp_effects_CEWL_best_mod1.csv")
```










