---
title: "Experimental Data Analysis"
author: "Savannah Weaver"
date: "June 2021"
output: 
  pdf_document:
    toc: TRUE
---

# Packages

```{r setup, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse")
if (!require("UsingR")) install.packages("UsingR")
library("UsingR")
if (!require("PerformanceAnalytics")) install.packages("PerformanceAnalytics") 
library("PerformanceAnalytics") # pretty multicollinearity plots
if (!require("MASS")) install.packages("MASS")
library("MASS") # model selection & box-cox
if (!require("lme4")) install.packages("lme4")
library("lme4") # for LMMs
if (!require("ggpubr")) install.packages("ggpubr")
library("ggpubr") # for Brown-Forsythe test
```


# Data

## Morphometrics & Hydration

### Treatment Groups

variables:
- individual lizard ID
- temp_tmt_C = temperature treatment
- humidity_tmt_percent = humidity treatment (high/low, not actually %)
- trial_number = which set of lizards that individual was from
- conclusion = how that individual's experiment ended (died, canceled, or complete)

```{r tmt data}
tmts <- read.csv("./data/exp_tmt_assignment.csv")
```


### Capture Data

variables:
- date = date of capture & baseline measurements
- individual lizard ID
- mass_g = mass in grams
- hematocrit_percent = % of blood sample that's red blood cells
- osmolality_mmol_kg = concentration of solutes in blood plasma
- type = when the measurements were taken along the course of the experiment (all on capture day)

```{r capture data}
capture_hydration <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  mutate(# correctly format date-only variable
         date = as.Date(date, format = "%Y-%m-%d")
         ) %>%
  # select only relevant variables
  dplyr::select(date, individual_ID, 
                mass_g, hematocrit_percent, osmolality_mmol_kg
                ) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID) %>%
  mutate(type = as.factor("capture"))
summary(capture_hydration)
```


extract SVL data separately from capture data:

```{r SVL data}
SVL <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::select(individual_ID, SVL_mm) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID)
summary(SVL)
```


extract capture CEWL cloacal temperature separately:

```{r cap CT data}
cap_CT <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::select(individual_ID, cloacal_temp_C) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID)
summary(cap_CT)
```



### Experiment Data

variables:
- date = date of measurements
- individual lizard ID
- mass_g = mass in grams
- hematocrit_percent = % of blood sample that's red blood cells
- osmolality_mmol_kg = concentration of solutes in blood plasma
- type = when the measurements were taken along the course of the experiment (either during experimental treatment or after rehab)

```{r exp data}
exp_dat <- read.csv("./data/experimental_data.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
                # format date
  dplyr::mutate(date = as.Date(date, format = "%m/%d/%y"),
                type = as.factor(type)
                ) %>%
  # select only variables to be analyzed
  dplyr::select(date, individual_ID, mass_g, 
                hematocrit_percent, type, 
                osmolality_mmol_kg = osmolality_mmol_kg_replicate_mean)
summary(exp_dat)
```

### Join Dataframes

Now, attach all the dataframes, only use individuals whose treatment was completed, and add a "day" variable for what day of treatment each lizard/observation was on. I also calculate SMI using the equation created in capture_analysis.


```{r join morpho blood data}
all_dat <- exp_dat %>%
  # join data
  rbind(capture_hydration) %>%
  # add tmt group info
  left_join(tmts, by = "individual_ID") %>%
  dplyr::select(-notes) %>%
  # add SVL value for each obs of each indiv.
  # for computing BCI and scaled mass indices
  left_join(SVL, by = "individual_ID") %>%
  # only use completed experiment runs
  dplyr::filter(conclusion == "complete") %>%
  group_by(individual_ID) %>%
  # reformat a lot of variables
  mutate(capture_date = min(date),
         day = as.numeric(date - capture_date),
         humidity_tmt_percent = as.factor(humidity_tmt_percent),
         individual_ID = as.factor(individual_ID),
         temp_tmt_C = as.factor(temp_tmt_C),
         trial_number = as.factor(trial_number),
         conclusion = as.factor(conclusion),
         SMI = mass_g * ((65.02158/SVL_mm) ^ (3.09059/sqrt(0.8944)))
         )

summary(all_dat)
```

re-order some factors:

```{r reorder all_dat humidity factor}
all_dat$humidity_tmt_percent <- factor(all_dat$humidity_tmt_percent,
                                       levels = c("humid", "dry"),
                                       labels = c("Humid", "Dry"))
```


make a sub-dataframe without rehab data to prevent any mix-ups:

```{r extra df}
all_dat_no_rehab <- all_dat %>%
  dplyr::filter(type != "rehab")
```


### Checks

Dates:

```{r check dates}
# check that capture dates are valid
unique(all_dat$capture_date)
```

Check that each lizard only has an accurate number of measurements.

```{r check n measures per lizard}
all_dat %>%
  group_by(individual_ID, type) %>%
  summarise(n = n()) %>%
  arrange(type)
```


That all looks good, experimental measurements are either 4 (first trial) or 2 (other trials). I am excluding lizards that died in treatment from the analysis.


## CEWL

### Capture CEWL

variables:
- date = date of capture & baseline measurements
- individual lizard ID
- region = which body area the measurement was taken from
- TEWL_g_m2h = evaporative water loss
- cloacal_temp_C = taken at measurement; influences CEWL

```{r cap CEWL data}
cap_CEWL <- read.csv("./exported_data/capture_CEWL.csv") %>%
  dplyr::select(date, individual_ID, region, TEWL_g_m2h) %>%
  mutate(#individual_ID = as.factor(individual_ID), # do later
         date = as.Date(date, format = "%Y-%m-%d"),
         region = as.factor(region),
         day = as.factor("before"), # might change... tbd
         n_day = 0
         ) %>%
  dplyr::filter(individual_ID %in% all_dat$individual_ID) %>%
  left_join(cap_CT, by = 'individual_ID')
summary(cap_CEWL)
```


### Post-Experiment CEWL

*In the future, I could automate this like I did for the HOBO data.*

Load in each of the post-rehab datafiles:

```{r exp CEWL data}
# trial 1
CEWL_t1 <- read.csv("./data/post_exp_CEWL/4-28-21-CEWL.csv", # filename
                          na.strings=c("","NA")) %>% # fix empty cells
  # rename and select the pertinent variables/cols
  # I have to do this for each one 
  # so they all have the same number of columns for joining
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.. # rename
                )

# trial 2
CEWL_t2 <- read.csv("./data/post_exp_CEWL/5-4-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )

# trial 3
CEWL_t3 <- read.csv("./data/post_exp_CEWL/5-11-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )

# trial 4
CEWL_t4 <- read.csv("./data/post_exp_CEWL/5-18-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )
```


Load in cloacal temperatures:

```{r exp CT data}
exp_CT <- read.csv("./data/post_exp_CEWL_cloacal_temps.csv") %>%
  mutate(date = as.Date(date, format = "%Y/%m/%d")) %>%
  dplyr::select(-time)
summary(exp_CT)
```


### Join Dataframes 

Merge all post-experiment CEWL, add cloacal temperature, add capture CEWL:

```{r join CEWL data}
# merge all CEWL datafiles & reformat
CEWL <- CEWL_t1 %>% # trial 1
  rbind(., CEWL_t2, # trial 2
        CEWL_t3, # trial 3
        CEWL_t4 # trial 4
        ) %>%
  # remove any unsuccessful measurements
  dplyr::filter(Status == "Normal") %>%
  # extract individual_ID and region separately from the "ID" variable
  separate(ID, c("individual_ID", "region")) %>%
  # reformat data
  dplyr::mutate(# reformat date
                date = as.Date(date, format = "%m/%d/%y"),
                # format individual ID as a factor
                individual_ID = as.integer(individual_ID),
                # set body region as a factor variable after getting only the consistent characters due to typos
                region = as.factor(substring(region, 1, 4)),
                # add when measurement taken
                day = as.factor("after"), # may rename...
                n_day = 1
                ) %>%
  # remove cols not relevant to stats
  dplyr::select(-Status) %>%
  # remove any rows with missing values
  dplyr::filter(complete.cases(.)) %>%
  # add cloacal temperatures
  left_join(exp_CT, by = c("date", "individual_ID")) %>%
  # now matching dataframes, add capture CEWL data
  rbind(cap_CEWL) %>%
  # add tmt assignments
  left_join(tmts, by = "individual_ID") %>%
  mutate(humidity_tmt_percent = as.factor(humidity_tmt_percent),
         individual_ID = as.factor(individual_ID),
         conclusion = as.factor(conclusion),
         trial_number = as.factor(trial_number)
         ) 
# every lizard should have 10 measurements
summary(CEWL)
```

Before/after aren't perfectly even because sometimes we were unable to get the AquaFlux to equilibrate and take a measurement.

Finally, make a small edit so the regions are spelled out completely. This requires reordering factor levels:

```{r reorder CEWL factors}
CEWL$region <- factor(CEWL$region, 
                      levels = c("head", "dewl", "dors", "vent", "mite"),
                      labels = c("Head", "Dewlap", "Dorsum", 
                                 "Ventrum", "Mite Patch")
                  )
CEWL$humidity_tmt_percent <- factor(CEWL$humidity_tmt_percent,
                                       levels = c("humid", "dry"),
                                       labels = c("Humid", "Dry"))
```



## Export Data Frames for Power Analyses

```{r write data}
write.csv(all_dat, "exported_data/exp_effects_hydration.csv")
write.csv(CEWL, "exported_data/exp_effects_CEWL.csv")
```





# Data Distributions

## Histograms

### Mass

```{r mass hist}
all_dat %>%
  ggplot(., aes(x = mass_g)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Mass (g)") + 
  ylab("Count")
simple.eda(all_dat$mass_g)
shapiro.test(all_dat$mass_g)
```

Mass distribution not normal, skewed to the left.


### Scaled Mass Index

```{r SMI hist}
all_dat %>%
  ggplot(., aes(x = SMI)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Scaled Mass Index (g)") + 
  ylab("Count")
simple.eda(all_dat$SMI)
shapiro.test(all_dat$SMI)
```


### Hematocrit

```{r hct hist}
all_dat %>%
  ggplot(., aes(x = hematocrit_percent)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Hematocrit (%)") + 
  ylab("Count")
simple.eda(all_dat$hematocrit_percent)
shapiro.test(all_dat$hematocrit_percent)
```

Visually, looks slightly skewed to the right, but statistically, the distribution of hematocrit is normal.


### Osmolality

```{r osml hist}
all_dat %>%
  ggplot(., aes(x = osmolality_mmol_kg)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Osmolality (mmol / kg)") + 
  ylab("Count")
simple.eda(all_dat$osmolality_mmol_kg)
shapiro.test(all_dat$osmolality_mmol_kg)
```

Visually, looks slightly skewed to the right, but statistically, the distribution of osmolality is normal.










# Basic Figures & Models

## Mass ~ Time

I won't be using this, SMI is more applicable.

Just look at plot:

```{r mass fig}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = mass_g, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = mass_g, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  xlab("Day") + 
  ylab("Mass (g)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0
)
```



## SMI ~ Time

plot over course of experiment:

```{r SMI fig}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = SMI, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = SMI, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  geom_line(aes(x = day,
                y = SMI,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  xlab("") + 
  ylab("Scaled Mass Index (g)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 18),
        legend.text.align = 0,
        legend.position = "none"
) -> tmt_effects_SMI
tmt_effects_SMI

# export figure
ggsave(filename = "tmt_effects_SMI.jpeg",
       plot = tmt_effects_SMI,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 5, height = 4)
```

Check whether means started out different:

```{r SMI check original difference}
SMI_diff_lm <- all_dat_no_rehab %>%
  dplyr::filter(day == 0) %>%
  lm(data = ., SMI ~humidity_tmt_percent)
summary(SMI_diff_lm)
```

NOT significantly different, which is good. t-test and p-value have very similar results

model:

```{r SMI models}
SMI_mod <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ day*humidity_tmt_percent +
               (1|trial_number))
summary(SMI_mod)
drop1(SMI_mod)

# drop interaction term
SMI_mod2 <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ day + humidity_tmt_percent +
               (1|trial_number))
summary(SMI_mod2)
drop1(SMI_mod2)
```

SMI is best predicted by day and treatment, but not including their interaction.

```{r}
write.csv(data.frame((summary(SMI_mod2))$coefficients), 
          "./best models/exp_effects_SMI.csv")
```


## Hct ~ Time

```{r hct fig}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = hematocrit_percent, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = hematocrit_percent, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  geom_line(aes(x = day,
                y = hematocrit_percent,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  xlab("Day") + 
  ylab("Hematocrit (%)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text.align = 0,
        legend.position = "none"
        ) -> tmt_effects_hct
tmt_effects_hct

# export figure
ggsave(filename = "tmt_effects_hct.jpeg",
       plot = tmt_effects_hct,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 5, height = 4)
```

this model seemed to work well with indiv as a random factor, but still excluded because it's probably unnecessary

```{r hct models}
hct_mod <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ day + humidity_tmt_percent +
               (1|trial_number))
summary(hct_mod)
drop1(hct_mod)

# drop humidity
hct_mod2 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ day +
               (1|trial_number))
summary(hct_mod2)
drop1(hct_mod2)
```

The model AIC is slightly better without the interaction effect, so I removed that. The effect of humidity could ALSO be dropped, so humidity treatment was not an important factor affecting hematocrit, but how many days lizards were in treatment was. Both treatment groups lost hematocrit at approximately the same rate. 

```{r}
write.csv(data.frame((summary(hct_mod2))$coefficients), 
          "./best models/exp_effects_hct.csv")
```



## Osml ~ Time

```{r osmolality fig}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = osmolality_mmol_kg, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = osmolality_mmol_kg, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  geom_line(aes(x = day,
                y = osmolality_mmol_kg,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  xlab("") + 
  ylab("Osmolality (mmol / kg)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text.align = 0,
        legend.position = "none"
        ) -> tmt_effects_osml
tmt_effects_osml

# export figure
ggsave(filename = "tmt_effects_osml.jpeg",
       plot = tmt_effects_osml,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 5, height = 4)
```

singular warning - do NOT include individual ID as a random effect

```{r osmolality model}
osml_mod <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>%
  lme4::lmer(data = .,
               osmolality_mmol_kg ~ day * humidity_tmt_percent +
               (1|trial_number))
summary(osml_mod)
drop1(osml_mod)
```

The model seems good as-is. 

```{r}
write.csv(data.frame((summary(osml_mod))$coefficients), 
          "./best models/exp_effects_osml.csv")
```



## Change in Osmolality

```{r osmol change data}
osml_d0 <- all_dat_no_rehab %>%
  dplyr::filter(day == 0) %>%
  dplyr::select(individual_ID, osml0 = osmolality_mmol_kg, 
                humidity_tmt_percent)
osml_d8 <- all_dat_no_rehab %>%
  dplyr::filter(day %in% c(8,9)) %>%
  dplyr::select(individual_ID, osml89 = osmolality_mmol_kg, 
                humidity_tmt_percent)

osml_diffs <- osml_d0 %>%
  left_join(osml_d8) %>%
  mutate(osml_change = osml89 - osml0)
```


boxplot:

```{r osmol change fig}
osml_diffs %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = humidity_tmt_percent, 
                   y = osml_change, 
                   group = humidity_tmt_percent,
                   color = humidity_tmt_percent
                   ), 
               size = 1,
               alpha = 1) + 
  theme_classic() + 
  xlab("") + 
  ylab("Osmolality Change (mmol / kg)") + 
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```

model to test statistical significance of difference:

```{r osmol change model}
osml_diff_mod <- lm(data = osml_diffs,
                    osml_change ~ humidity_tmt_percent)
summary(osml_diff_mod)
```
Not difference between the two treatment groups.


## CEWL ~ Before/After

try a boxplot:

```{r CEWL fig 1}
CEWL %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = humidity_tmt_percent, 
                   y = TEWL_g_m2h, 
                   color = day
                   ), 
               size = 1,
               alpha = 1) + 
  facet_wrap(~region) +
  theme_classic() + 
  xlab("") + 
  ylab("CEWL (g / m^2h)") + 
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "bottom"
)
```

this is difficult to see changes, I think a line graph would be better...

```{r CEWL fig 2}
CEWL %>% 
  ggplot(data = .) + 
  geom_point(aes(x = n_day,
                 y = TEWL_g_m2h, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  geom_line(aes(x = n_day,
                y = TEWL_g_m2h,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  stat_smooth(aes(x = n_day, 
                  y = TEWL_g_m2h, 
                  color = humidity_tmt_percent
                  ),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1) + 
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  facet_wrap(~region, ncol = 2) +
  scale_x_continuous(breaks = c(0, 1),
                     labels = c("0" = "before", "1" = "after")
                     ) +
  xlab("") + 
  ylab(bquote('CEWL (g/'*m^2*'h)')) + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10,
                                 angle = 90),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
        legend.position = c(0.75,0.12),
        #legend.justification = c(1, 1)
) -> CEWL_tmt_fig
CEWL_tmt_fig

# export figure
ggsave(filename = "tmt_effects_CEWL.jpeg",
       plot = CEWL_tmt_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 4, height = 6)
```

I saved the legend separately to make the figure layout better.

*TO DO: figure out how to annotate legend onto figure?!*

```{r CEWL model 1}
CEWL_mod <- CEWL %>%
  dplyr::filter(complete.cases(.)) %>%
  lme4::lmer(data = .,
               TEWL_g_m2h ~ day * humidity_tmt_percent * region +
               (1|trial_number/individual_ID))
summary(CEWL_mod)
drop1(CEWL_mod)
```

Drop triple interaction. I think the day:region standalone would be weird too.

```{r CEWL model 2}
CEWL_mod2 <- CEWL %>%
  dplyr::filter(complete.cases(.)) %>%
  lme4::lmer(data = .,
               TEWL_g_m2h ~ 
               day*humidity_tmt_percent +
               humidity_tmt_percent*region +
               (1|trial_number/individual_ID))
summary(CEWL_mod2)
drop1(CEWL_mod2)
```

We can drop the humidity:region interaction.

```{r CEWL model 3}
CEWL_mod3 <- CEWL %>%
  dplyr::filter(complete.cases(.)) %>%
  lme4::lmer(data = .,
               TEWL_g_m2h ~ 
               day*humidity_tmt_percent + region +
               (1|trial_number/individual_ID))
summary(CEWL_mod3)
drop1(CEWL_mod3)
```

The model is best with all the parameters currently included in model 3. 

```{r}
write.csv(data.frame((summary(CEWL_mod3))$coefficients), 
          "./best models/exp_effects_CEWL.csv")
```


## Change in CEWL

```{r CEWL change data}
CEWL_before <- CEWL %>%
  dplyr::filter(day == "before") %>%
  dplyr::select(CEWL_before = TEWL_g_m2h,
                humidity_tmt_percent, trial_number,
                individual_ID, region)
CEWL_after <- CEWL %>%
  dplyr::filter(day == "after") %>%
  dplyr::select(CEWL_after = TEWL_g_m2h,
                humidity_tmt_percent, trial_number,
                individual_ID, region)

CEWL_diffs <- CEWL_before %>%
  left_join(CEWL_after, by = c('individual_ID', 'region',
                'humidity_tmt_percent', 'trial_number')) %>%
  mutate(CEWL_diff = CEWL_after - CEWL_before)
```


plot:

```{r CEWL change fig}
CEWL_diffs %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = humidity_tmt_percent, 
                   y = CEWL_diff, 
                   group = humidity_tmt_percent,
                   color = humidity_tmt_percent
                   ), 
               size = 1,
               alpha = 1) + 
  #facet_wrap(~humidity_tmt_percent) +
  theme_classic() + 
  geom_hline(yintercept = 0, lty = 2) +
  xlab("") + 
  ylab("CEWL Change (g/m2h)") + 
  #annotate("text", x = 1.5, y = 45, 
   #        label = "paste(italic(p), \" = 0.0152\")", 
    #       parse = TRUE,
     #      size = 6) +
  #ylim(10, 50) +
  #scale_x_discrete(labels = c("F" = "Female",
   #                           "M" = "Male")) +
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```


model:

```{r CEWL change model}
CEW_diffs_mod <- lm(data = CEWL_diffs,
                    CEWL_diff ~ humidity_tmt_percent)
summary(CEW_diffs_mod)
```


## Multi-Figure

```{r}
ggarrange(tmt_effects_SMI, tmt_effects_osml, tmt_effects_hct,
          ncol = 1, nrow = 3,
          common.legend = TRUE,
          legend = "bottom"
          ) -> tmt_multi_fig
tmt_multi_fig
# export figure
ggsave(filename = "tmt_multi_fig.jpeg",
       plot = tmt_multi_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 12)
```

# Rehydration...

## Data

First, get only the data for before experiment, after experiment, and after rehab.

```{r rehydration model}
summary(all_dat)
rehydrat_dat <- all_dat %>%
  dplyr::filter(day %in% c(0, 8, 9, 10, 11))
rehydrat_dat$day <- factor(rehydrat_dat$day,
                           levels = c(0, 8, 9, 10, 11),
                           labels = c("Before Experiment",
                                      "After Experiment",
                                      "After Experiment",
                                      "After Rehydration",
                                      "After Rehydration"))
summary(rehydrat_dat)
```

## SMI

```{r rehydration SMI fig}
rehydrat_dat %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = SMI, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 1) +
  geom_line(aes(x = day,
                y = SMI,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.6) +
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  scale_x_discrete(labels = c("Before\nExperiment",
                              "After\nExperiment",
                              "After\nRehydration")) +
  xlab("") + 
  xlab("") + 
  ylab("Scaled Mass Index (g)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "bottom"
        ) -> rehab_SMI_fig
rehab_SMI_fig

# export figure
ggsave(filename = "rehab_SMI_fig.jpeg",
       plot = rehab_SMI_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 5, height = 4)
```


## Osmolality

first, make a list of all the IDs that have a post-rehab osmolality measurement, since this has a lot of missing data

```{r rehydration osmol data}
rehab_osmols <- rehydrat_dat %>%
  dplyr::filter(day == "After Rehydration") %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg))
```


```{r rehydration osmol fig}
rehydrat_dat %>% 
  dplyr::filter(individual_ID %in% rehab_osmols$individual_ID) %>%
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = osmolality_mmol_kg, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 1) +
  geom_line(aes(x = day,
                y = osmolality_mmol_kg,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.6) +
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  scale_x_discrete(labels = c("Before\nExperiment",
                              "After\nExperiment",
                              "After\nRehydration")) +
  xlab("") + 
  xlab("") + 
  ylab("Osmolality (mmol/kg)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "bottom"
  ) -> rehab_osml_fig
rehab_osml_fig

# export figure
ggsave(filename = "rehab_osml_fig.jpeg",
       plot = rehab_osml_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 5, height = 4)
```



## Hematocrit

first, make a list of all the IDs that have all three measurements:

```{r rehydration hct data}
rehab_hct <- rehydrat_dat %>%
  dplyr::filter(day == "After Rehydration") %>%
  dplyr::filter(complete.cases(hematocrit_percent))
```


```{r rehydration hct fig}
rehydrat_dat %>% 
  dplyr::filter(individual_ID %in% rehab_osmols$individual_ID) %>%
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = hematocrit_percent, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 1) +
  geom_line(aes(x = day,
                y = hematocrit_percent,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.6) +
  theme_classic() + 
  scale_color_brewer(palette = "Set2",
                     name = "Humidity Treatment") +
  scale_x_discrete(labels = c("Before\nExperiment",
                              "After\nExperiment",
                              "After\nRehydration")) +
  xlab("") + 
  ylab("Hematocrit (%)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 12),
        legend.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "bottom"
  ) -> rehab_hct_fig
rehab_hct_fig

# export figure
ggsave(filename = "rehab_hct_fig.jpeg",
       plot = rehab_hct_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 5, height = 4)
```


## Multi-Figure

```{r}
ggarrange(rehab_SMI_fig, rehab_osml_fig, rehab_hct_fig,
          ncol = 1, nrow = 3,
          common.legend = TRUE,
          legend = "bottom"
          ) -> rehab_multi_fig
rehab_multi_fig
# export figure
ggsave(filename = "rehab_multi_fig.jpeg",
       plot = rehab_multi_fig,
       path = "./final_figures",
       device = "jpeg",
       dpi = 1200,
       width = 6, height = 12)
```



