---
title: "Experimental Data Analysis"
author: "Savannah Weaver"
date: "June 2021"
output: 
  github_document:
    toc: TRUE
---

# Packages

```{r setup, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
library("tidyverse")
if (!require("UsingR")) install.packages("UsingR")
library("UsingR")
if (!require("PerformanceAnalytics")) install.packages("PerformanceAnalytics") 
library("PerformanceAnalytics") # pretty multicollinearity plots
if (!require("MASS")) install.packages("MASS")
library("MASS") # model selection & box-cox
if (!require("lme4")) install.packages("lme4")
library("lme4") # for LMMs
```


# Data

## Morphometrics & Hydration

### Treatment Groups

variables:
- individual lizard ID
- temp_tmt_C = temperature treatment
- humidity_tmt_percent = humidity treatment (high/low, not actually %)
- trial_number = which set of lizards that individual was from
- conclusion = how that individual's experiment ended (died, canceled, or complete)

```{r}
tmts <- read.csv("./data/exp_tmt_assignment.csv")
```


### Capture Data

variables:
- date = date of capture & baseline measurements
- individual lizard ID
- mass_g = mass in grams
- hematocrit_percent = % of blood sample that's red blood cells
- osmolality_mmol_kg = concentration of solutes in blood plasma
- type = when the measurements were taken along the course of the experiment (all on capture day)

```{r capture data}
capture_hydration <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  mutate(# correctly format date-only variable
         date = as.Date(date, format = "%Y-%m-%d")
         ) %>%
  # select only relevant variables
  dplyr::select(date, individual_ID, 
                mass_g, hematocrit_percent, osmolality_mmol_kg
                ) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID) %>%
  mutate(type = as.factor("capture"))
summary(capture_hydration)
```


extract SVL data separately from capture data:

```{r SVL}
SVL <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::select(individual_ID, SVL_mm) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID)
summary(SVL)
```


extract capture CEWL cloacal temperature separately:

```{r cap CT}
cap_CT <- read.csv("./exported_data/capture_hydration.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
  dplyr::select(individual_ID, cloacal_temp_C) %>%
  dplyr::filter(individual_ID %in% tmts$individual_ID)
summary(cap_CT)
```



### Experiment Data

variables:
- date = date of measurements
- individual lizard ID
- mass_g = mass in grams
- hematocrit_percent = % of blood sample that's red blood cells
- osmolality_mmol_kg = concentration of solutes in blood plasma
- type = when the measurements were taken along the course of the experiment (either during experimental treatment or after rehab)

```{r}
exp_dat <- read.csv("./data/experimental_data.csv",
                             na.strings=c("","NA") # fix empty cells
                             ) %>%
                # format date
  dplyr::mutate(date = as.Date(date, format = "%m/%d/%y"),
                type = as.factor(type)
                ) %>%
  # select only variables to be analyzed
  dplyr::select(date, individual_ID, mass_g, 
                hematocrit_percent, type, 
                osmolality_mmol_kg = osmolality_mmol_kg_replicate_mean)
summary(exp_dat)
```

### Join Dataframes

Now, attach all the dataframes, only use individuals whose treatment was completed, and add a "day" variable for what day of treatment each lizard/observation was on. I also calculate SMI using the equation created in capture_analysis.


```{r}
all_dat <- exp_dat %>%
  # join data
  rbind(capture_hydration) %>%
  # add tmt group info
  left_join(tmts, by = "individual_ID") %>%
  dplyr::select(-notes) %>%
  # add SVL value for each obs of each indiv.
  # for computing BCI and scaled mass indices
  left_join(SVL, by = "individual_ID") %>%
  # only use completed experiment runs
  dplyr::filter(conclusion == "complete") %>%
  group_by(individual_ID) %>%
  # reformat a lot of variables
  mutate(capture_date = min(date),
         day = as.numeric(date - capture_date),
         humidity_tmt_percent = as.factor(humidity_tmt_percent),
         individual_ID = as.factor(individual_ID),
         temp_tmt_C = as.factor(temp_tmt_C),
         trial_number = as.factor(trial_number),
         conclusion = as.factor(conclusion),
         SMI = mass_g * ((65.02158/SVL_mm) ^ (3.09059/sqrt(0.8944)))
         )

summary(all_dat)
```

make a sub-dataframe without rehab data to prevent any mix-ups:

```{r}
all_dat_no_rehab <- all_dat %>%
  dplyr::filter(type != "rehab")
```


### Checks

Dates:

```{r}
# check that capture dates are valid
unique(all_dat$capture_date)
```

Check that each lizard only has one measurement for each day of measurements.

```{r}
all_dat %>%
  group_by(individual_ID, type) %>%
  summarise(n = n()) %>%
  arrange(type)
```


That all looks good, experimental measurements are either 4 (first trial) or 2 (other trials). I am excluding lizards that died in treatment from the analysis.


## CEWL

### Capture CEWL

variables:
- date = date of capture & baseline measurements
- individual lizard ID
- region = which body area the measurement was taken from
- TEWL_g_m2h = evaporative water loss
- cloacal_temp_C = taken at measurement; influences CEWL

```{r}
cap_CEWL <- read.csv("./exported_data/capture_CEWL.csv") %>%
  dplyr::select(date, individual_ID, region, TEWL_g_m2h) %>%
  mutate(#individual_ID = as.factor(individual_ID), # do later
         date = as.Date(date, format = "%Y-%m-%d"),
         region = as.factor(region),
         day = as.factor("before"), # might change... tbd
         n_day = 0
         ) %>%
  dplyr::filter(individual_ID %in% all_dat$individual_ID) %>%
  left_join(cap_CT, by = 'individual_ID')
summary(cap_CEWL)
```


### Post-Experiment CEWL

*In the future, I could automate this like I did for the HOBO data.*

Load in each of the post-rehab datafiles:

```{r}
# trial 1
CEWL_t1 <- read.csv("./data/post_exp_CEWL/4-28-21-CEWL.csv", # filename
                          na.strings=c("","NA")) %>% # fix empty cells
  # rename and select the pertinent variables/cols
  # I have to do this for each one 
  # so they all have the same number of columns for joining
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h.. # rename
                )

# trial 2
CEWL_t2 <- read.csv("./data/post_exp_CEWL/5-4-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )

# trial 3
CEWL_t3 <- read.csv("./data/post_exp_CEWL/5-11-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )

# trial 4
CEWL_t4 <- read.csv("./data/post_exp_CEWL/5-18-21-CEWL.csv",
                          na.strings=c("","NA")) %>%
  dplyr::select(date = Date, 
                Status,
                ID = Comments,
                TEWL_g_m2h = TEWL..g..m2h..
                )
```


Load in cloacal temperatures:

```{r}
exp_CT <- read.csv("./data/post_exp_CEWL_cloacal_temps.csv") %>%
  mutate(date = as.Date(date, format = "%Y/%m/%d")) %>%
  dplyr::select(-time)
summary(exp_CT)
```


### Join Dataframes 

Merge all post-experiment CEWL, add cloacal temperature, add capture CEWL:

```{r}
# merge all CEWL datafiles & reformat
CEWL <- CEWL_t1 %>% # trial 1
  rbind(., CEWL_t2, # trial 2
        CEWL_t3, # trial 3
        CEWL_t4 # trial 4
        ) %>%
  # remove any unsuccessful measurements
  dplyr::filter(Status == "Normal") %>%
  # extract individual_ID and region separately from the "ID" variable
  separate(ID, c("individual_ID", "region")) %>%
  # reformat data
  dplyr::mutate(# reformat date
                date = as.Date(date, format = "%m/%d/%y"),
                # format individual ID as a factor
                individual_ID = as.integer(individual_ID),
                # set body region as a factor variable after getting only the consistent characters due to typos
                region = as.factor(substring(region, 1, 4)),
                # add when measurement taken
                day = as.factor("after"), # may rename...
                n_day = 1
                ) %>%
  # remove cols not relevant to stats
  dplyr::select(-Status) %>%
  # remove any rows with missing values
  dplyr::filter(complete.cases(.)) %>%
  # add cloacal temperatures
  left_join(exp_CT, by = c("date", "individual_ID")) %>%
  # now matching dataframes, add capture CEWL data
  rbind(cap_CEWL) %>%
  # add tmt assignments
  left_join(tmts, by = "individual_ID") %>%
  mutate(humidity_tmt_percent = as.factor(humidity_tmt_percent),
         individual_ID = as.factor(individual_ID),
         conclusion = as.factor(conclusion),
         trial_number = as.factor(trial_number)
         ) 
# every lizard should have 10 measurements
summary(CEWL)
```

Before/after aren't perfectly even because sometimes we were unable to get the AquaFlux to equilibrate and take a measurement post-experiment.


## Export Data Frames for Power Analyses

```{r}
write.csv(all_dat, "exported_data/exp_effects_hydration.csv")
write.csv(CEWL, "exported_data/exp_effects_CEWL.csv")
```





# Data Distributions

## Histograms

### Mass

```{r}
exp_dat %>%
  ggplot(., aes(x = mass_g)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Mass (g)") + 
  ylab("Count")
simple.eda(exp_dat$mass_g)
shapiro.test(exp_dat$mass_g)
```

Mass distribution not normal, skewed to the left.


### Body Condition Index

```{r}
exp_dat %>%
  ggplot(., aes(x = mass_g)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Mass (g)") + 
  ylab("Count")
simple.eda(exp_dat$mass_g)
shapiro.test(exp_dat$mass_g)
```


### Hematocrit

```{r}
exp_dat %>%
  ggplot(., aes(x = hematocrit_percent)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Hematocrit (%)") + 
  ylab("Count")
simple.eda(exp_dat$hematocrit_percent)
shapiro.test(exp_dat$hematocrit_percent)
```

Visually, looks slightly skewed to the right, but statistically, the distribution of hematocrit is normal.


### Osmolality

```{r}
exp_dat %>%
  ggplot(., aes(x = osmolality_mmol_kg)) +
  geom_histogram(color = "black", fill="steelblue", bins=50) + 
  theme_classic() +
  xlab("Osmolality (mmol / kg)") + 
  ylab("Count")
simple.eda(exp_dat$osmolality_mmol_kg)
shapiro.test(exp_dat$osmolality_mmol_kg)
```

Visually, looks slightly skewed to the right, but statistically, the distribution of osmolality is normal.










# Basic Figures & SLR

### Mass ~ Time

I won't be using this, SMI is more applicable.

Just look at plot:

```{r}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = mass_g, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = mass_g, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  xlab("Day") + 
  ylab("Mass (g)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0
)
```



### SMI ~ Time

plot over course of experiment:

```{r}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = SMI, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = SMI, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  geom_line(aes(x = day,
                y = SMI,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  xlab("Day") + 
  ylab("Scaled Mass Index (g)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0
)
```

Check whether means started out different:

```{r}
t.test()
```


model:

```{r}
SMI_mod <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ day*humidity_tmt_percent +
               (1|trial_number))
summary(SMI_mod)
drop1(SMI_mod)

# drop interaction term
SMI_mod2 <- lme4::lmer(data = all_dat_no_rehab,
               SMI ~ day + humidity_tmt_percent +
               (1|trial_number))
summary(SMI_mod2)
drop1(SMI_mod2)
```

SMI is best predicted by day and treatment, but not including their interaction.

boxplot comparing before exp, after exp, and after rehab:

```{r}

```



### Hct ~ Time

```{r}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = hematocrit_percent, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = hematocrit_percent, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  geom_line(aes(x = day,
                y = hematocrit_percent,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  xlab("Day") + 
  ylab("Hematocrit (%)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0
)
```

this model seemed to work well with indiv as a random factor, but still excluded because it's probably unnecessary

```{r}
hct_mod <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ day + humidity_tmt_percent +
               (1|trial_number))
summary(hct_mod)
drop1(hct_mod)

# drop humidity
hct_mod2 <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(hematocrit_percent)) %>%
  lme4::lmer(data = .,
               hematocrit_percent ~ day +
               (1|trial_number))
summary(hct_mod2)
drop1(hct_mod2)
```

The model AIC is slightly better without the interaction effect, so I removed that. The effect of humidity could ALSO be dropped, so humidity treatment was not an important factor affecting hematocrit, but how many days lizards were in treatment was. Both treatment groups lost hematocrit at approximately the same rate. 

boxplot comparing before exp, after exp, and after rehab:

```{r}

```



### Osml ~ Time

```{r}
all_dat_no_rehab %>% 
  ggplot(data = .) + 
  geom_point(aes(x = day,
                 y = osmolality_mmol_kg, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  stat_smooth(aes(x = day, 
                  y = osmolality_mmol_kg, 
                  color = humidity_tmt_percent
                  ), 
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1 ) + 
  geom_line(aes(x = day,
                y = osmolality_mmol_kg,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  theme_classic() + 
  scale_x_continuous(breaks = c(0, 2, 4, 6, 8)) +
  xlab("Date") + 
  ylab("Osmolality (mmol / kg)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0
)
```

singular warning - do NOT include individual ID as a random effect

```{r}
osml_mod <- all_dat_no_rehab %>%
  dplyr::filter(complete.cases(osmolality_mmol_kg)) %>%
  lme4::lmer(data = .,
               osmolality_mmol_kg ~ day * humidity_tmt_percent +
               (1|trial_number))
summary(osml_mod)
drop1(osml_mod)
```

The model seems good as-is. 


boxplot comparing before exp, after exp, and after rehab:

```{r}

```



### Change in Osmolality

```{r}
osml_d0 <- all_dat_no_rehab %>%
  dplyr::filter(day == 0) %>%
  dplyr::select(individual_ID, osml0 = osmolality_mmol_kg, 
                humidity_tmt_percent)
osml_d8 <- all_dat_no_rehab %>%
  dplyr::filter(day %in% c(8,9)) %>%
  dplyr::select(individual_ID, osml89 = osmolality_mmol_kg, 
                humidity_tmt_percent)

osml_diffs <- osml_d0 %>%
  left_join(osml_d8) %>%
  mutate(osml_change = osml89 - osml0)
```


boxplot:

```{r}
osml_diffs %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = humidity_tmt_percent, 
                   y = osml_change, 
                   group = humidity_tmt_percent
                   ), 
               size = 1,
               alpha = 1) + 
  theme_classic() + 
  xlab("") + 
  ylab("Osmolality Change (mmol / kg)") + 
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```

model to test statistical significance of difference:

```{r}
osml_diff_mod <- lm(data = osml_diffs,
                    osml_change ~ humidity_tmt_percent)
summary(osml_diff_mod)
```
Not different between the two treatment groups.


### CEWL ~ Before/After

try a boxplot:

```{r}
CEWL %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = humidity_tmt_percent, 
                   y = TEWL_g_m2h, 
                   color = day
                   ), 
               size = 1,
               alpha = 1) + 
  facet_wrap(~region) +
  theme_classic() + 
  xlab("") + 
  ylab("CEWL (g / m^2h)") + 
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "bottom"
)
```

this is difficult to see changes, I think a line graph would be better...

```{r}
CEWL %>% 
  ggplot(data = .) + 
  geom_point(aes(x = n_day,
                 y = TEWL_g_m2h, 
                 color = humidity_tmt_percent
                 ), 
             size = 1, 
             alpha = 0.6) +
  geom_line(aes(x = n_day,
                y = TEWL_g_m2h,
                group = individual_ID,
                color = humidity_tmt_percent),
            alpha = 0.2) +
  stat_smooth(aes(x = n_day, 
                  y = TEWL_g_m2h, 
                  color = humidity_tmt_percent
                  ),
              formula = y ~ x, 
              method = "lm", 
              se = F, 
              size = 1.6, 
              alpha = 1) + 
  theme_classic() + 
  facet_wrap(~region) +
  scale_x_continuous(breaks = c(0, 1),
                     labels = c("0" = "before", "1" = "after")
                     ) +
  xlab("") + 
  ylab("TEWL (g / m2h)") + 
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 12),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 10),
        legend.text.align = 0,
        legend.position = "bottom"
)
```

Okay, to-do: rotate x-axis labels, redo legend

```{r}
CEWL_mod <- CEWL %>%
  dplyr::filter(complete.cases(.)) %>%
  lme4::lmer(data = .,
               TEWL_g_m2h ~ day * humidity_tmt_percent * region +
               (1|trial_number/individual_ID))
summary(CEWL_mod)
drop1(CEWL_mod)
```

Drop triple interaction. I think the day:region standalone would be weird too.

```{r}
CEWL_mod2 <- CEWL %>%
  dplyr::filter(complete.cases(.)) %>%
  lme4::lmer(data = .,
               TEWL_g_m2h ~ 
               day*humidity_tmt_percent +
               humidity_tmt_percent*region +
               (1|trial_number/individual_ID))
summary(CEWL_mod2)
drop1(CEWL_mod2)
```

We can drop the humidity:region interaction.

```{r}
CEWL_mod3 <- CEWL %>%
  dplyr::filter(complete.cases(.)) %>%
  lme4::lmer(data = .,
               TEWL_g_m2h ~ 
               day*humidity_tmt_percent + region +
               (1|trial_number/individual_ID))
summary(CEWL_mod3)
drop1(CEWL_mod3)
```

The model is best with all the parameters currently included in model 3. 


### Change in CEWL

```{r}
CEWL_before <- CEWL %>%
  dplyr::filter(day == "before") %>%
  dplyr::select(CEWL_before = TEWL_g_m2h,
                humidity_tmt_percent, trial_number,
                individual_ID, region)
CEWL_after <- CEWL %>%
  dplyr::filter(day == "after") %>%
  dplyr::select(CEWL_after = TEWL_g_m2h,
                humidity_tmt_percent, trial_number,
                individual_ID, region)

CEWL_diffs <- CEWL_before %>%
  left_join(CEWL_after, by = c('individual_ID', 'region',
                'humidity_tmt_percent', 'trial_number')) %>%
  mutate(CEWL_diff = CEWL_after - CEWL_before)
```


plot:

```{r}
CEWL_diffs %>% 
  ggplot(data = .) + 
  geom_boxplot(aes(x = humidity_tmt_percent, 
                   y = CEWL_diff, 
                   group = humidity_tmt_percent,
                   color = humidity_tmt_percent
                   ), 
               size = 1,
               alpha = 1) + 
  #facet_wrap(~humidity_tmt_percent) +
  theme_classic() + 
  geom_hline(yintercept = 0, lty = 2) +
  xlab("") + 
  ylab("CEWL Change (g/m2h)") + 
  #annotate("text", x = 1.5, y = 45, 
   #        label = "paste(italic(p), \" = 0.0152\")", 
    #       parse = TRUE,
     #      size = 6) +
  #ylim(10, 50) +
  #scale_x_discrete(labels = c("F" = "Female",
   #                           "M" = "Male")) +
  scale_color_brewer(palette = "Set2") +
  theme(text = element_text(color = "black", 
                            family = "sans", 
                            size = 18),
        axis.text = element_text(color = "black", 
                                 family = "sans", 
                                 size = 14),
        legend.text.align = 0,
        legend.position = "none"
)
```


model:

```{r}
CEW_diffs_mod <- lm(data = CEWL_diffs,
                    CEWL_diff ~ humidity_tmt_percent)
summary(CEW_diffs_mod)
```




# Rehydration... still need to do!





