---
title: "HOBO Logger Data for Sceloporus Climate Studies"
author: "Savannah Weaver"
output: pdf_document
toc: TRUE
---

# Packages

```{r setup, echo=T, results='hide', message=FALSE}
if (!require("tidyverse")) install.packages("tidyverse") 
library("tidyverse")
```

# Load in Data

This data was collected using Onset HOBO temperature and humidity dataloggers during the course of our experiment. See __ for full details. 

The data is in a separate file for each download for each logger, so I need to compile each of those into one dataset. 

To do this, first I compile a list of the filenames I need to read-in.

```{r, filenames}
# make a list of file names of all data to load in
filenames <- list.files(path = "data/HOBOs")
```

Next, I make a function that will read in the data from each csv, name and organize the data correctly. 

```{r, create read_HOBO_files function to get each csv data}
# make a function to read in data from each csv file and add correct identifiers
read_HOBO_files <- function(filename) {
  
  # edit the filename inputted to funtion
  # to make a unique identifier for each logger
  name <- substr(filename, 1, nchar(filename)-15)
  
  # also make an identifier for the site
  site <- substr(filename, 1, nchar(filename)-15)
  
  # read in the csv file for this given filename
  dat <- read.csv(file.path("data/HOBOs", filename),
                # each csv has headers
                header = TRUE,
                # this is what I want to rename the col headers
                col.names = c("order", "date_time_PST", "temp_C", 
                              "relative_humidity", "dew_pt_C", 
                              # the 6,7,8th cols are not data
                              # logger use info we don't need
                              "mostly_blank", "mostly_blank", "mostly_blank")
                ) %>%
    # select only the cols with data we want
    # don't need order- just an arbitrary observation identifier
    # don't need "mostly_blank" cols- unnecessary logger use info
    # but get the rest of the cols with informative data
    dplyr::select(date_time_PST, temp_C, relative_humidity, dew_pt_C) %>%
    # add a column with the name of the HOBO the data is from 
    dplyr::mutate(HOBO_ID = name)
  
  # return the dataframe for that single csv file
  dat
}
```

Finally, I apply the function I made to all of the filenames I compiled, then put all of those dataframes into one dataframe for my analyses.

This will print warnings saying that header and col.names are different lengths, because the data has extra notes on logger usage that we read-in, but get rid of.

```{r, apply read_HOBO_files function}
# apply function to get data from all csvs
all_HOBO_data <- lapply(filenames, read_HOBO_files) %>%
  # paste all data files together into one df by row
  reduce(rbind)
```



# Data Wrangling

Now that I have a full dataframe, I want to make sure all the data is there and correct. 

```{r}
str(all_HOBO_data)
summary(all_HOBO_data)
```

I need to subset the data based on a date range to extract data for actual experiment days, since the loggers run continuously. There are also NA's which are an artifact of the logger's notes.

```{r}
exp_dates <- as.Date(c("2021-04-19", "2021-04-20", 
                       "2021-04-21", "2021-04-22",
               "2021-04-23", "2021-04-24", "2021-04-25", "2021-04-26",
               "2021-04-27", "2021-04-28", "2021-04-29", "2021-04-30",
               "2021-05-01", "2021-05-02", "2021-05-03", "2021-05-04",
               "2021-05-05", "2021-05-06", "2021-05-07", "2021-05-08",
               "2021-05-09", "2021-05-10", "2021-05-11", "2021-05-12",
               "2021-05-13", "2021-05-14", "2021-05-15", "2021-05-16",
               "2021-05-17", "2021-05-18"), 
               format = "%Y-%m-%d")

subset_HOBO_data <- all_HOBO_data %>%
  dplyr::filter(complete.cases(temp_C, relative_humidity)) %>%
  mutate(date_only = as.Date(date_time_PST, format = "%m/%d/%Y")) %>%
  dplyr::filter(date_only %in% exp_dates)
```

Next, format the data:

```{r}
format_HOBO_data <- subset_HOBO_data %>% 
  mutate(date_time_PST = as.POSIXct(date_time_PST, 
                                    format = "%m/%d/%Y %H:%M:%S"),
         HOBO_ID = as.factor(substr(HOBO_ID, 1, 10))
         )
summary(format_HOBO_data)
```

I'm also going to omit the data for the hot treatments that failed, since we will not be presenting that.

Check temperature values:

```{r}
format_HOBO_data %>% 
  ggplot(data = ., aes(x = temp_C)) +
  geom_histogram()
format_HOBO_data %>%
  group_by(date_only, HOBO_ID) %>%
  summarise(mean_temp = mean(temp_C)) %>%
  arrange(mean_temp, date_only)
```

We do NOT want data for Dry-Hot-1/2 or Humid-Hot between April 19-29. Everything else should be fine, so just filter those out.

The goal is to get everything EXCEPT those three loggers on those dates. The most straightforward way I know how to do this is to just filter out by temperature.

**NOT WORKING**
```{r}
filter_dates <- as.Date(c("2021-04-19", "2021-04-20", 
                       "2021-04-21", "2021-04-22",
               "2021-04-23", "2021-04-24", "2021-04-25", "2021-04-26",
               "2021-04-27", "2021-04-28", "2021-04-29"), 
               format = "%Y-%m-%d")
filter_HOBOs <- c("Dry-Hot-1", "Dry-Hot-2", "Humid-Hot")
unique(format_HOBO_data$HOBO_ID)
# get subsets
filt_keep1 <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID == 'Dry-Hot-1')
filt_keep2 <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID %in% c("Dry-Hot-1", "Dry-Hot-2", "Humid-Hot") & 
                  date_only %nin% filter_dates)
filt_dontwant_check <- format_HOBO_data %>%
  dplyr::filter(HOBO_ID %in% c("Dry-Hot-1", "Dry-Hot-2", "Humid-Hot") & 
                  date_only %in% filter_dates)

format_HOBO_data[format_HOBO_data$HOBO_ID == ('Dry-Hot-1')]

# rejoin
filtered_HOBO_data <- format_HOBO_data %>%
  dplyr::filter(temp_C < 28)

# check that it worked correctly
filtered_HOBO_data %in% filt_dontwant_check
```


**NOT WORKING**
```{r}
levels(format_HOBO_data$HOBO_ID) <- list("Dry" = "Dry-Cool-1",
                                         "Dry-Cool-2" = "Humid",
                                         "Dry-Hot-1" = "Dry",
                                         "Dry-Hot-2" = "Humid",
                                         "Humid-Cool" = "Humid",
                                         "Humid-Hot" = "Humid")
                                         
  format_HOBO_data$HOBO_ID,	29.54950	6.492429	64.37734	32.061846
	25.58470	1.865721	84.76560	14.486066

                                   levels = c("Dry1", "Humid1", "Dry2", 
                                              "Humid2", "Humid3", "Humid4"),
                                   labels = c("Dry1", "Humid1", "Dry2", 
                                              "Humid2", "Humid3", "Humid4"))
```


re-check:

```{r}
summary(format_HOBO_data)
```


# Get Means

```{r}
format_HOBO_data %>%
  group_by(HOBO_ID) %>%
  summarise(mean_temp_C = mean(temp_C),
            sd_temp_C = sd(temp_C),
            mean_humidity = mean(relative_humidity),
            sd_humidity = sd(relative_humidity))
```

















